---
title: "Olympic Data"
author: "team 010100"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  rmdformats::readthedown:
    css: styles.css
    highlight: kate
    code_folding: hide
    number_sections: yes
    keep_tex: yes
    includes:
        after_body: github.html
bibliography: bibliography.bib
csl: american-institute-of-physics.csl
---

```{r basic, include=F}
# use this function to conveniently load libraries and work smoothly with knitting
# can add quietly=T option to the require() function
loadPkg = function(pkg, character.only = FALSE) { 
  if (!character.only) { pkg <- as.character(substitute(pkg)) }
  pkg <- ifelse(!character.only, as.character(substitute(pkg)) , pkg)  
  if (!require(pkg,character.only=T, quietly =T)) {  install.packages(substitute(pkg),dep=T); if(!require(pkg,character.only=T)) stop("Package not found") } 
}
loadPkg(knitr)

# unload/detact package when done using it
unloadPkg = function(pkg, character.only = FALSE) { 
  if(!character.only) { pkg <- as.character(substitute(pkg)) } 
  search_item <- paste("package", pkg,sep = ":") 
  while(search_item %in% search()) { detach(search_item, unload = TRUE, character.only = TRUE) } 
}
```


```{r setup, echo=FALSE, cache=FALSE}
loadPkg(knitr)
loadPkg(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```


# Introduction
Team 010100 are the following members: Izzy Illari, Lucia Illari, Omar Qusous, and Lydia Teinfalt. You may find our work over on [GitHub](https://github.com/IzzyIllari/DATS-Model-Proj).

For the second portion of our group project, we kept Olympics data from the EDA. Our SMART questions were `What factors can be used to model the probability of being awarded a medal?` What groups/clusters do athletes of different sports fall into? How does a pandemic affect the medals awarded? How can the evolution of athlete characteristics over time be modelled? With these questions in mind we went to see if we could find use the data on Olympians to find patterns and create models that could answer the questions.

We used a dataset called `120 years of Olympic history: athletes and results` on Kaggle over here: [https://www.kaggle.com/heesoo37/120-years-of-olympic-history-athletes-and-results](https://www.kaggle.com/heesoo37/120-years-of-olympic-history-athletes-and-results). This historical dataset includes all Olympic Games from Athens 1896 to Rio 2016, which was scraped from [https://www.sports-reference.com/](https://www.sports-reference.com/). We focused on data from Olympic events 1960-2016 when looking at clustering, Kmeans, Linear and Logit Regression and trends over time. For the pandemic analysis, we focused on data of Olympics participating in events before and after the H1N1 Pandemic from 1918-1919.   

The report is organized as follows:

1. Summary of Dataset
2. Data Prep
3. EDA
4. Clustering, Kmeans, Kmedoids 
5. Linear and Logit Regression
6. Random Forest
7. Pandemic (Spanish Flu)
8. Trends over time
9. Summary and Conclusion
10. References

# Summary of Dataset
The data looks like the following:

```{r xkablesummary, include = FALSE}
loadPkg(xtable)
loadPkg(kableExtra)
loadPkg(stringi)

xkabledply = function(smmry, title='Caption', pos='left') { # Thanks Ryan Longmuir for the codes
  smmry %>%
    xtable() %>% 
    kable(caption = title, digits = 4) %>%
    kable_styling(position = "center") %>%
    kable_styling(bootstrap_options = "striped", full_width = F,
    position = pos)
}

xkablesummary = function(df) { 
  #' Combining base::summary, xtable, and kableExtra, to easily display numeric variable summary of dataframes. 
  #` If the categorical variables has less than 6 levels, the function will still run without error.
  #' ELo 202003 GWU DATS
  #' version 1
  #' @param df The dataframe.
  #' @return The summary table for display, or for knitr to process into other formats 
  #' @examples
  #' xkablesummary( faraway::ozone )
  #' xkablesummary( ISLR::Hitters )
  
  s = summary(df) %>%
    apply( 2, function(x) stringr::str_remove_all(x,c("Min.\\s*:\\s*","1st Qu.\\s*:\\s*","Median\\s*:\\s*","Mean\\s*:\\s*","3rd Qu.\\s*:\\s*","Max.\\s*:\\s*")) ) %>% # replace all leading words
    apply( 2, function(x) stringr::str_trim(x, "right")) # trim trailing spaces left
  
  colnames(s) <- stringr::str_trim(colnames(s))
  
  if ( dim(s)[1] ==6 ) { rownames(s) <- c('Min','Q1','Median','Mean','Q3','Max') 
  } else if ( dim(s)[1] ==7 ) { rownames(s) <- c('Min','Q1','Median','Mean','Q3','Max','NA') }
  
  s %>%
    xkabledply("Table: Statistics summary.", "center")

}
```

```{r import_data}
olympic_data <- data.frame(read.csv("olympic_data.csv"))
olympic_data$ID <- as.factor(olympic_data$ID)
olympic_data$BMI.Category <- as.factor(olympic_data$BMI.Category)
olympic_data$Medal.No.Yes <- as.factor(olympic_data$Medal.No.Yes)
#head(olympic_data)
str(olympic_data)
```


The athlete events data has `r ncol(olympic_data)` columns and `r nrow(olympic_data)` rows/entries, for a total of `r ncol(olympic_data)*nrow(olympic_data)` individual data points. In `olympic_data` each row corresponds to an individual athlete competing in an individual Olympic event. The variables are the following:

1. ID: Unique number for each athlete
2. Name: Athlete's name
3. Sex: M or F
4. Age: Integer
5. Height: centimeters
6. Weight: kilograms
7. Team: Team name
8. NOC: National Olympic Committee 3-letter code
9. Games: Year and season
10. Year: Integer
11. Season: Summer or Winter
12. City: Host city
13. Sport
14. Event
15. Medal: Gold, Silver, Bronze, or NA

To prepare our data for EDA we dropped the Olympic event: Art Sculpting. NAs were also removed. We have modified the data from the kaggle dataset from which it was originally taken. The dataset now starts at 1960 and includes the new following variables:

1. Decade (factor)
2. First name (factor)
3. Last name (factor)
4. BMI (numeric)
5. BMI category (factor)
6. Population (numeric)
7. GDP (numeric)
8. GDPpC (numeric)
9. Medal: Yes or No (factor)


```{r remove_NAs, echo = FALSE}
olympic.data <- olympic_data

loadPkg(VIM)
aggr(olympic.data)
unloadPkg(VIM)

olympic_data_noNA <- na.omit(olympic_data)
olympic_data <- olympic_data_noNA
data_sport <- split(olympic_data_noNA, olympic_data_noNA$Sport)
```


# EDA

For EDA, we can do a quick summary to just look at the data.

```{r keep, results = "markup"}
xkablesummary(olympic_data)

# loadPkg(psych)
# pairs.panels(olympic_data[,-length(olympic_data)], 
#              method = "pearson",
#              hist.col = "#CCFF66",
#              density = TRUE,
#              ellipses = TRUE
#              )
# unloadPkg(psych)

olympic_data$Sex.Int <- c(as.numeric(as.factor(olympic_data$Sex)))
olympic_data$NOC.Int <- c(as.numeric(as.factor(olympic_data$NOC)))
olympic_data$Sport.Int <- c(as.numeric(as.factor(olympic_data$Sport)))
```

# Olympics Correlation plot

Just quickly visualizing thw correlation will be useful for model building, but we have to be mindful of the fact that columns such as Medal and Medal.No.Yes are noturally going to be highly correlated.

```{r subset_corr, echo = FALSE}
loadPkg("dplyr")
olympics_subset <- olympic_data %>% filter(!is.na(Age)) %>% select(Year, NOC.Int, Sex.Int, Age, Height, Weight, BMI, BMI.Category, Population, GDP, GDPpC, Medal.No.Yes)
unloadPkg("dplyr")

loadPkg("corrplot")
cols.num <- c(1:length(olympic.data))
num.df <- olympic.data
num.df[cols.num] <- sapply(olympic.data[cols.num],as.numeric)
cor.all <- cor(num.df[,c(1,2,8:13,15:17,24)], use="pairwise.complete.obs")
cmat.all <- corrplot(cor.all, method="pie", type = "upper", order = "hclust", tl.col = "black", tl.srt = 45)
unloadPkg("corrplot")
```

It might be more useful to focus in on the correlations for only the variable Medal.No.Yes and Medal:

```{r focus_corr, results = "markup"}
loadPkg(corrr)
loadPkg(plyr)

cor.medNY <- focus(correlate(num.df, use="pairwise.complete.obs"), Medal.No.Yes)
cor.medNY[order(cor.medNY$Medal.No.Yes, decreasing = TRUE),] %>%
  kable("html", align = 'cccc') %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),full_width = T)

cor.medal <- focus(correlate(num.df, use="pairwise.complete.obs"), Medal)
cor.medal[order(cor.medal$Medal, decreasing = TRUE),] %>%
  kable("html", align = 'cccc') %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),full_width = T)

unloadPkg(corrr)
unloadPkg(plyr)
```

Unless you ignore the athletes that didn't receive a medal, building a general model from the variable Medal.No.Yes might be a better idea, based off the strength of the correlations.

# Clustering, Kmeans, Kmedoids 
# Logit Regression
# KNN
# Random Forest
# Pandemic (Spanish Flu)
With the novel corona virus pandemic this year and its delay of the Olympics in Tokyo this year, we thought it would be interesting to study a last pandemic in the last century and analyze the impact it had on the Olympic performance. The following countries in Europe had 2.64 million excess deaths occurred during the period when the H1N1 Pandemic (also commonly called Spanish Flu) was circulating from January 1918 - June 1919: Italy, Bulgaria, Portugal, Spain, Netherlands, Sweden, Germany, Switzerland, France, Norway, Denmark, UK (Scotland, England, Wales). In the US, 675,000 people died from H1N1 which was 0.8 percent of the 1910 population. 

```{r loadpackages}
loadPkg(digest)
loadPkg(corrplot)
loadPkg(ggcorrplot)
loadPkg(ggplot2)
loadPkg(rmdformats)
loadPkg(knitr)
loadPkg(lattice) 
loadPkg(jtools)
loadPkg(faraway)
loadPkg(leaps)
loadPkg(gridExtra)
```

```{r read_pandemic_data}
p_olympics <- read.csv("pandemic_olympics.csv")
```
(JOHNSON, NIALL P. A. S., and JUERGEN MUELLER. “Updating the Accounts: Global Mortality of the 1918-1920 ‘Spanish’ Influenza Pandemic.” Bulletin of the History of Medicine, vol. 76, no. 1, 2002, pp. 105–115. JSTOR, www.jstor.org/stable/44446153. Accessed 19 Apr. 2020.) Taken from [@Johnson2002]

Of the European countries that suffered significant excess deaths during the Spanish Influenza Pandemic, these countries competed before and after 1918-1919: Denmark (DEN), France (FRA), Great Britain (GBR), Italy (ITA), Netherlands (NED), Norway (NOR), Sweden (SWE), and United States (USA). We created a separate pandemic data set containing athletes from these countries that competed in the Olympics between 1908-1928 to study before and after the pandemic. 

```{r H1N1_pandemic}
loadPkg(dplyr)
NOC_SF <- c("ITA", "NED", "SWE", "FRA", "NOR", "DEN", "GBR", "USA")
Medals <- c("Gold", "Silver", "Bronze")

pandemic_NOC_Yr_Mdl <- p_olympics %>% filter(Year >= 1908 & Year <= 1928,
                                         NOC %in% NOC_SF,
                                         Medal %in% Medals) %>% group_by(NOC, Year) %>% tally()

pandemic_NOC_Yr_Mdl$Year <- as.factor(pandemic_NOC_Yr_Mdl$Year)
unloadPkg(dplyr)

```
## Medals Earned 
The plot shows number of medals earned by Denmark (DEN), France (FRA), Great Britain (GBR), Italy (ITA), Netherlands (NED), Norway (NOR), Sweden (SWE), and United States (USA) before and after the pandemic. There may be more than Gold, Silver or Bronze medals earned by individuals in any sporting event; accounting for the group events.
```{r plots_num_medals , include=TRUE}
NOC_colors <- c("#A6CEE3", "#1F78B4", "#B2DF8A", "#33A02C", "#FB9A99", "#E31A1C", "#FDBF6F","#FF7F00")


pandemic_df <- data.frame(pandemic_NOC_Yr_Mdl$NOC, pandemic_NOC_Yr_Mdl$Year, pandemic_NOC_Yr_Mdl$n)
pandemic_df <- unique(pandemic_df)
colnames(pandemic_df) <- c("NOC", "Year", "Total.Medals")

pandemic_df$NOC <- as.factor(pandemic_df$NOC)


# Basic line plot with points
ggplot(data=pandemic_df, aes(x=Year, y=Total.Medals, group=NOC, color=NOC)) +
  geom_line(size=1)+
#  scale_colour_manual(values = NOC_colors) +
   geom_vline(xintercept = 1918) +
  geom_point()+
 theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, size = 9, hjust = 1),
        axis.text.y = element_text(vjust = 1, size = 9),
        panel.border = element_rect(color = "black", fill=NA, size=1)) +
  labs(title="H1N1 Pademic: Num of medals (n) vs. Year (1908-1928)", x="Year", y="Number of medals (n)") 


```
Great Britain (GBR), Denmark (DEN), Sweden (SWE) saw a decline in the number of medals their athletes earned after the pandemic. The Olympics were not held in 1916 due to World War I. 

## Number of Olympic Athletes
Looking at the same countries, the plot shows the number of athletes they sent to the Olympics from 1908 - 1928. 
```{r plots_num_athletes, include=TRUE}
loadPkg("dplyr")
NOC_SF <- c("ITA", "NED", "SWE", "FRA", "NOR", "DEN", "GBR", "USA")

p_num_athletes <- p_olympics %>% filter(Year >= 1908 & Year <= 1928,
                                         NOC %in% NOC_SF) %>% group_by(NOC, Year) %>% tally(unique(ID))


p_num_athletes_df <- data.frame(p_num_athletes$NOC, p_num_athletes$Year, p_num_athletes$n)
p_num_athletes_df <- unique(p_num_athletes_df)
colnames(p_num_athletes_df) <- c("NOC", "Year", "Total.Athletes")
p_num_athletes_df$NOC <- as.factor(p_num_athletes_df$NOC)
p_num_athletes_df$Year <- as.factor(p_num_athletes_df$Year)
p_num_athletes_df$V_line <- c("1918") 


# Basic line plot with points
ggplot(data=p_num_athletes_df, aes(x=Year, y=Total.Athletes, group=NOC, color=NOC)) +
  geom_line(size=1)+
#  scale_colour_manual(values = NOC_colors) +
   geom_vline(xintercept = 1918) +
  geom_point()+
 theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, size = 9, hjust = 1),
        axis.text.y = element_text(vjust = 1, size = 9),
        panel.border = element_rect(color = "black", fill=NA, size=1)) +
  labs(title="H1N1 Pademic: Num of athletes (n) vs. Year (1908-1928)", x="Year", y="Number of Athletes (n)")
  
unloadPkg("dplyr")
```
Great Britain and Sweden saw a sharp decline in the number of athletes that they sent to Olympic events after the pandemic. JOHNSON, NIALL P. A. S., and JUERGEN MUELLER reported England & Wales had approximately 200,000 death toll (per 1,000) and Sweden reported 34,374 death toll during the 1918-1919 pandemic. 

## Average Age of Olympians
The chart displays the average age of Olympians from the eight countries in our data before and after the pandemic.  
```{r ageSF}
loadPkg(dplyr)
NOC_colors <- c("#A6CEE3", "#1F78B4", "#B2DF8A", "#33A02C", "#FB9A99", "#E31A1C", "#FDBF6F","#FF7F00")

pandemic_avg_age <- p_olympics %>% filter(Year >= 1908 & Year <= 1928,
                                         NOC %in% NOC_SF,
                                         !is.na(Age)) %>% group_by(NOC, Year) %>% summarise(avg=mean(Age))

pandemic_avg_age$Year <- as.factor(pandemic_avg_age$Year)


# Basic line plot with points
ggplot(data=pandemic_avg_age, aes(x=Year, y=avg, group=NOC, color=NOC)) +
  geom_line(size=1)+
   geom_vline(xintercept = 1918) +
  geom_point()+
 theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, size = 9, hjust = 1),
        axis.text.y = element_text(vjust = 1, size = 9),
        panel.border = element_rect(color = "black", fill=NA, size=1)) +
  labs(title="H1N1 Pademic: Average Age vs. Year (1908-1928)", x="Year", y="Avg Age of Athletes (n)")
```
The H1N1 Influenza pandemic "Spanish flu" was fatal for individuals aged 20–40 years. The average age of Olympians competing after the pandemic was increased for all countries in our data set.  

## Average Height and Weight of Olympians
```{r w_h}
pandemic_avg_hw<- p_olympics %>% filter(Year >= 1908 & Year <= 1928,
                                         NOC %in% NOC_SF,
                                        !is.na(Height),
                                         !is.na(Weight)) %>% group_by(NOC, Year) %>% summarise(avg_h = mean(Height), avg_w = mean(Weight))
pandemic_avg_hw$Year <- as.factor(pandemic_avg_hw$Year)


# Basic line plot with points
h1 <- ggplot(data=pandemic_avg_hw, aes(x=Year, y=avg_h, group=NOC, color=NOC)) +
  geom_line(size=1)+
   geom_vline(xintercept = 1918) +
  geom_point()+
 theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, size = 9, hjust = 1),
        axis.text.y = element_text(vjust = 1, size = 9),
        panel.border = element_rect(color = "black", fill=NA, size=1)) +
  labs(title="H1N1 Pademic: Average Height vs. Year (1908-1928)", x="Year", y="Avg Height (cm)")
h1

w1 <- ggplot(data=pandemic_avg_hw, aes(x=Year, y=avg_w, group=NOC, color=NOC)) +
  geom_line(size=1)+
   geom_vline(xintercept = 1918) +
  geom_point()+
 theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, size = 9, hjust = 1),
        axis.text.y = element_text(vjust = 1, size = 9),
        panel.border = element_rect(color = "black", fill=NA, size=1)) +
  labs(title="H1N1 Pademic: Average Weight vs. Year (1908-1928)", x="Year", y="Avg Height (cm)")
w1

```
Netherlands (NED) and Norway (NOR) had significant increase in the average height and weight of their athletes after Spanish flu pandemic. Sweden (SWE) and France (FRA) saw a decrease in those averages. 


```{r model1}
loadPkg(dplyr)
olympics_subset <- olympic_data %>% filter(!is.na(Age)) %>% select(Year, NOC.Int, Sex.Int, Age, Height, Weight, BMI, BMI.Category, Population, GDP, GDPpC, Medal.No.Yes)

olympics_medal_count <- olympics_subset %>% group_by(Year) %>% tally(as.numeric(Medal.No.Yes))

p_olympics$Medal.No.Yes <- ifelse(p_olympics$Medal == "Gold" | p_olympics$Medal == "Silver" | p_olympics$Medal == "Bronze",1,0)
p_olympics$Medal.No.Yes[is.na(p_olympics$Medal.No.Yes)] <- 0

#p_olympics$Medal.No.Yes <- as.factor(p_olympics$Medal.No.Yes)


pandemic_total_medals_gold<- p_olympics %>% filter(Medal == "Gold", Season == "Summer") %>% group_by(NOC, Year) %>% tally(Medal.No.Yes)
#pandemic_total_medals_gold
pandemic_total_medals_silver<- p_olympics %>% filter(Medal == "Silver", Season == "Summer") %>% group_by(NOC, Year) %>% tally(Medal.No.Yes)
#pandemic_total_medals_silver
pandemic_total_medals_bronze<- p_olympics %>% filter(Medal == "Bronze", Season == "Summer") %>% group_by(NOC, Year) %>% tally(Medal.No.Yes)
#pandemic_total_medals_bronze
pandemic_total_medals <- p_olympics %>% filter(Medal %in% Medals, Season == "Summer") %>% group_by(NOC, Year) %>% tally(Medal.No.Yes)
#pandemic_total_medals

df<- data.frame(p_olympics$Medal, p_olympics$Medal.No.Yes)


unloadPkg(dplyr)
```

## Total Number of Olympic Medals (Summer Events) vs. Year 
```{r }
loadPkg(dplyr)

olympics_total_medals <- olympic_data %>% filter(Medal %in% Medals, Season == "Summer") %>%
group_by(NOC, Year) %>% tally(as.numeric(Medal.No.Yes))
#olympics_total_medals

olympics_summer_medals <- olympic_data %>% filter(Medal %in% Medals, Season == "Summer") %>%
group_by(Year) %>% tally(as.numeric(Medal.No.Yes))
colnames(olympics_summer_medals) <- c("Year", "Medals")
olympics_summer_medals$Year <- as.factor(olympics_summer_medals$Year)



ITA_Tmedals <- pandemic_total_medals %>% filter(NOC == "ITA")  
#ITA_Tmedals
ITA_Tmedals_all <- olympics_total_medals  %>% filter(NOC == "ITA")  
#ITA_Tmedals_all


#histogram of Number of Medals
loadPkg(ggplot2)
ggplot(data=olympics_summer_medals, aes(x=Year, y=Medals)) +
  geom_bar(stat="identity", fill="steelblue")+ylab("Total Number of Medals (Summer Events)") + theme_minimal()

unloadPkg(dplyr)
```

## Creating Time Series - Italy
During the 1918 Pandemic, Italy's death toll (per 1,000) was 390,000 death toll. According to Worldodometer,  (https://www.worldometers.info/coronavirus/country/italy/), during the current Covid-19 pandemic, there has been 29,079 deaths in Italy. We will focus on Italy to conduct time series analysis. Can we see a pattern with historical Olympic data before and after the Spanish flu in order to predict how Italy will fare in future Olympics after is emerges from the current novel Corona virus pandemic?

The Olympics data for Italy from 1908-1928 was converted to a time series and plotted "Total Number of Medals vs Year". 

```{r forecast}

olympic_pts <- ts(ITA_Tmedals$n, start=1908, end = 1928, deltat=4)
olympic_all <- ts(ITA_Tmedals_all$n, start=1908, end = 2016, deltat=4)

ita_olympic <- ts.union(olympic_pts, olympic_all)


loadPkg(tidyverse)
ita_df <- data.frame(ita_olympic)
ita_df <- ita_df %>% mutate(Num.Medals= if_else(is.na(ita_df$olympic_pts), ita_df$olympic_all,ita_df$olympic_pts ))

ita_df <- subset(ita_df, select = c(Num.Medals))


ol_ts <- ts(ita_df,start=1908, end=2016, deltat=4)
#plot(ol_ts)

autoplot(ol_ts, color="Blue") +
  ggtitle("Italy in the Olympics: Number of Medals") +
  xlab("Year") +
  ylab("Number of Medals")


olympic_train <- window(ol_ts, start=1908, end = 1928, deltat=4)

olympic_test <- window(ol_ts, start=2008, end = 2016, deltat=4)


acf(ol_ts,main="Italy Olympics")

```
The time series shows random fluctuations in the data over time, no overall trend. The Auto Correlated Function (ACF) plot does not show seasonality, periodicity and cyclic nature of the series. The ACF values are less than 0.05, so they are not significant. So the number of medals may not be correlated to time.


## Using 1908-1928 Time Series as Training Data
From the quick EDA, time series may not be appropriate method to model the number of medals Italy may earn in future Olympics. As a learning academic exercise, we will explore different time series methodologies to forecast. 
```{r forecast1}

olympic_pts <- ts(ITA_Tmedals$n, start=1908, end = 1928, deltat=4)
olympic_all <- ts(ITA_Tmedals_all$n, start=1960, end = 2016, deltat=4)

olympic_train <- window(olympic_pts, start=1908, end = 1928, deltat=4)
olympic_train1 <- window(ol_ts, start=1960, end = 2008, deltat=4)
olympic_test <- window(ol_ts, start=2008, end = 2016, deltat=4)

```

## Exploring Holt-Winters and ETS-ANN Time Series Forecasting 
Holt-Winters uses exponential smoothing to make short-term forecasts. Models is designated as either additive or multiplicative. 

```{r fit}
fit1.hw <- HoltWinters(x=olympic_train, beta = FALSE,gamma = FALSE) 
plot(fit1.hw)
summary(fit1.hw)

fit2.hw <- HoltWinters(x=olympic_train1, beta = FALSE, gamma = FALSE) 
plot(fit2.hw)
#summary(fit2.hw)


loadPkg(forecast)
fit1.ets <-ets(olympic_train1, model="ANN")
#plot(fit1.ets)
fit1_forecast <- forecast(fit1.ets, h = 3) 
fit1_forecast

autoplot(fit1_forecast)+
  ggtitle("ETS Decomposition: Number of Medals") +
  xlab("Year") +
  ylab("Number of Medals")


fit1.hw.forecast <- forecast(fit1.hw, 8)

autoplot(fit1.hw.forecast)+
  ggtitle("Holt-Winters Forecast 1908-28: Number of Medals") +
  xlab("Year") +
  ylab("Number of Medals")

fit2.hw.forecast <- forecast(fit2.hw, 8)

autoplot(fit2.hw.forecast)+
  ggtitle("Holt-Winters Forecast 1960-2008: Number of Medals") +
  xlab("Year") +
  ylab("Number of Medals")



```
```{r accuracy}
loadPkg(dplyr)
ITA_2016 <- olympics_total_medals %>% filter (NOC == "ITA" & Year == "2016")

```

## Time Series Linear Model for Italy
```{r }
h <- 10
fit.lin <- tslm(olympic_all ~ trend)
fcasts.lin <- forecast(fit.lin, h = h)
fit.exp <- tslm(olympic_all ~ trend, lambda = 0)
fcasts.exp <- forecast(fit.exp, h = h)

t <- time(olympic_all)
t.break1 <- 1908
t.break2 <- 1960
tb1 <- ts(pmax(0, t - t.break1), start = 1908)
tb2 <- ts(pmax(0, t - t.break2), start = 1908)

fit.pw <- tslm(olympic_all ~ t + tb1 + tb2)
t.new <- t[length(t)] + seq(h)
tb1.new <- tb1[length(tb1)] + seq(h)
tb2.new <- tb2[length(tb2)] + seq(h)

newdata <- cbind(t=t.new, tb1=tb1.new, tb2=tb2.new) %>%
  as.data.frame()
fcasts.pw <- forecast(fit.pw, newdata = newdata)

fit.spline <- tslm(olympic_all ~ t + I(t^2) + I(t^3) +
  I(tb1^3) + I(tb2^3))
fcasts.spl <- forecast(fit.spline, newdata = newdata)


autoplot(olympic_all) +
  autolayer(fitted(fit.lin), series = "Linear") +
  autolayer(fitted(fit.exp), series = "Exponential") +
  autolayer(fitted(fit.pw), series = "Piecewise") +
  autolayer(fitted(fit.spline), series = "Cubic Spline") +
  autolayer(fcasts.pw, series="Piecewise") +
  autolayer(fcasts.lin, series="Linear", PI=FALSE) +
  autolayer(fcasts.exp, series="Exponential", PI=FALSE) +
  autolayer(fcasts.spl, series="Cubic Spline", PI=FALSE) +
  xlab("Year") + ylab("Number of Medals") +
  ggtitle("Italy in the Olympics Time Series ") +
  guides(colour = guide_legend(title = " "))

```
Above plots try to fit the number of medals to an algorithm that could would best be able to predict the number of Olympic medals Italy will earn after Covid-19 pandemic. Visually, Cubic Spline seems closest in approximation.   

## Arima Model
```{r arima}
olympic_arima<- auto.arima(olympic_all, seasonal=FALSE)
summary(olympic_arima)

olympic_arima %>% forecast(h=10) %>% autoplot(include=8)

```

# Trends over time
# Summary and Conclusion
# References

