---
title: "Olympic Data"
author: "team 010100"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  rmdformats::readthedown:
    css: styles.css
    highlight: kate
    code_folding: hide
    number_sections: yes
    keep_tex: yes
    includes:
        after_body: github.html
bibliography: bibliography.bib
csl: american-institute-of-physics.csl
---

```{r basic, include=F}
# use this function to conveniently load libraries and work smoothly with knitting
# can add quietly=T option to the require() function
loadPkg = function(pkg, character.only = FALSE) { 
  if (!character.only) { pkg <- as.character(substitute(pkg)) }
  pkg <- ifelse(!character.only, as.character(substitute(pkg)) , pkg)  
  if (!require(pkg,character.only=T, quietly =T)) {  install.packages(substitute(pkg),dep=T); if(!require(pkg,character.only=T)) stop("Package not found") } 
}
loadPkg(knitr)

# unload/detact package when done using it
unloadPkg = function(pkg, character.only = FALSE) { 
  if(!character.only) { pkg <- as.character(substitute(pkg)) } 
  search_item <- paste("package", pkg,sep = ":") 
  while(search_item %in% search()) { detach(search_item, unload = TRUE, character.only = TRUE) } 
}
```


```{r setup, echo=FALSE, cache=FALSE}
loadPkg(knitr)
loadPkg(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

```{r import_data}
olympic_data <- data.frame(read.csv("olympic_data.csv"))
olympic_data$BMI.Category <- as.factor(olympic_data$BMI.Category)
olympic_data$Medal.No.Yes <- as.factor(olympic_data$Medal.No.Yes)
head(olympic_data)
str(olympic_data)
```

# Introduction
Team 010100 are the following members: Obumneke Amadi, Izzy Illari, Lucia Illari, Omar Qusous, and Lydia Teinfalt. You may find our work over on [GitHub](https://github.com/IzzyIllari/DATS-Model-Proj).

For the second portion of our group project, we kept Olympics data from the EDA. Our SMART questions were `What groups/clusters do athletes of different sports fall into?` What factors can be used to model the probability of being awarded a medal? How does a pandemic affect the medals awarded? How can the evolution of athlete characteristics over time be modelled? With these questions in mind we went to see if we could find use the data on Olympians to find patterns and create models that could answer the questions.

We used a dataset called `120 years of Olympic history: athletes and results` on Kaggle over here: [https://www.kaggle.com/heesoo37/120-years-of-olympic-history-athletes-and-results](https://www.kaggle.com/heesoo37/120-years-of-olympic-history-athletes-and-results). This historical dataset includes all Olympic Games from Athens 1896 to Rio 2016, which was scraped from [https://www.sports-reference.com/](https://www.sports-reference.com/). Except for the analysis of the H1N1 pandemic aka Spanish Flu, we focused on data from Olympic events 1960-2016.  

The report is organized as follows:

1. Summary of Dataset
2. Data Prep
3. EDA
4. Clustering, Kmeans, Kmedoids 
5. Linear and Logit Regression
6. Random Forest
7. Pandemic (Spanish Flu)
8. Trends over time
9. Summary and Conclusion
10. References

# Summary of Dataset
The data looks like the following:

```{r input_data}
athlete_events <- data.frame(read.csv("athlete_events.csv"))
#noc_regions <- data.frame(read.csv("noc_regions.csv"))
#summary(athlete_events)
str(athlete_events)
```

The athlete events data has `r ncol(athlete_events)` columns and `r nrow(athlete_events)` rows/entries, for a total of `r ncol(athlete_events)*nrow(athlete_events)` individual data points. In `athelete_events` each row corresponds to an individual athlete competing in an individual Olympic event. The variables are the following:

1. ID: Unique number for each athlete
2. Name: Athlete's name
3. Sex: M or F
4. Age: Integer
5. Height: centimeters
6. Weight: kilograms
7. Team: Team name
8. NOC: National Olympic Committee 3-letter code
9. Games: Year and season
10. Year: Integer
11. Season: Summer or Winter
12. City: Host city
13. Sport
14. Event
15. Medal: Gold, Silver, Bronze, or NA

To prepare our data for EDA we dropped the Olympic event: Art Sculpting. NAs were also removed. 

```{r remove_NAs}
athlete_events_noNA <- na.omit(athlete_events)
data_sport <- split(athlete_events_noNA, athlete_events_noNA$Sport)
```


# EDA
# Clustering, Kmeans, Kmedoids 
# Logit Regression
# KNN
# Random Forest
# Pandemic (Spanish Flu)
# Trends over time
# Summary and Conclusion
# References

