---
title: "Olympic Data"
author: "Izzy Illari"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  rmdformats::readthedown:
    css: styles.css
    highlight: kate
    code_folding: hide
    number_sections: yes
    keep_tex: yes
    includes:
        after_body: github.html
bibliography: bibliography.bib
csl: american-institute-of-physics.csl
---

```{r clear, include=F}
rm(list = ls(all.names = TRUE))
gc() 
```

```{r basic, include=F}
# use this function to conveniently load libraries and work smoothly with knitting
# can add quietly=T option to the require() function
loadPkg = function(pkg, character.only = FALSE) { 
  if (!character.only) { pkg <- as.character(substitute(pkg)) }
  pkg <- ifelse(!character.only, as.character(substitute(pkg)) , pkg)  
  if (!require(pkg,character.only=T, quietly =T)) {  install.packages(substitute(pkg),dep=T); if(!require(pkg,character.only=T)) stop("Package not found") } 
}
loadPkg(knitr)

# unload/detact package when done using it
unloadPkg = function(pkg, character.only = FALSE) { 
  if(!character.only) { pkg <- as.character(substitute(pkg)) } 
  search_item <- paste("package", pkg,sep = ":") 
  while(search_item %in% search()) { detach(search_item, unload = TRUE, character.only = TRUE) } 
}
```

```{r setup, echo=FALSE, cache=FALSE}
loadPkg(knitr)
loadPkg(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

# Import the data

```{r import_data}
olympic_data <- data.frame(read.csv("olympic_data.csv"))
olympic_data$BMI.Category <- as.factor(olympic_data$BMI.Category)
olympic_data$Medal.No.Yes <- as.factor(olympic_data$Medal.No.Yes)
head(olympic_data)
str(olympic_data)
```

```{r}
loadPkg(minpack.lm)
loadPkg(kableExtra)
loadPkg(ggplot2)
```

Number of sports per year at the https://www.topendsports.com/events/summer/sports/number.htm [@WoodOlympEvents]

```{r}
sports_dat <- data.frame(read.csv("sports_per_year.csv"))

ggplot(sports_dat, aes(x = Year, y=Num.Sports)) + 
  geom_point() + 
  geom_line() + 
  ylim(0, 40) + 
  theme_minimal() +
  labs(title="Plot of the number of sports at the Olympic Game over the years", 
       x="Year", y="Number of sports") + 
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1),
        axis.text.x = element_text(angle = 90, vjust = 0.5)) 
```


There is clearly an upward trend, but no seasonal pattern. The data is also a little choppy at the beginning. Part of the explanation is that the data points are not evenly spaced. Most Olympic games are 4 years apart, but a few of them are just 2 years apart, and during World War I and World War II there were 8-year and 12-year gaps, respectively. Since time series data should be evenly spaced over time, we'll only look at data from 1948 on, when the Olympics started being held every 4 years without any interruptions.

```{r}
index_1948 <- which(sports_dat$Year == 1948)
use_dat <- sports_dat[12:nrow(sports_dat),]
ggplot(use_dat, aes(x = Year, y=Num.Sports)) + 
  geom_point() + 
  geom_line() + 
  ylim(0, 40) + 
  theme_minimal() +
  labs(title="Plot of the number of sports at the Olympic Game over the years \n starting from 1948", 
       x="Year", y="Number of sports") + 
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1),
        axis.text.x = element_text(angle = 90, vjust = 0.5)) 
```

# Creating models

I'm going to try 4 different models.

$$
y_{\text{linear}}(x) = ax+b \\
y_{\text{quadratic}}(x) = ax^2 + bx + c \\
y_{\text{exponential}}(x) = a\exp(bx) + c \\
y_{\text{cubic}}(x) = ax^3 + bx^2 + cx + d 
$$

And I'll be able to use `ANOVA` to test the nested models: linear vs quadratic, and exponential growth vs s-curve (sigmoid). 

```{r}
lin_func <- function(x, a, b) {
  a*x+b
}

quad_func <- function(x, a, b, c) {
  a*x*x + b*x + c
}

cubic_func <- function(x, a, b, c, d) {
  a*x*x*x + b*x*x + c*x + d 
}

exp_func <- function(x, a, b, c) {
  a*exp(b*x) + c
}
```

Now I will try the model fits on the number of events per Olympic Games data. 

```{r}
xdata <- use_dat$Year
ydata <- use_dat$Num.Sports

mod_lin <- nlsLM(ydata~lin_func(xdata, a1, b1), start=list(a1=1, b1=10))
mod_quad <- nlsLM(ydata~quad_func(xdata, a1, b1, c1), start=list(a1=1, b1=1, c1=1))
mod_cubic <- nlsLM(ydata~cubic_func(xdata, a1, b1, c1, d1), start=list(a1=1, b1=1, c1=1, d1=1))
mod_exp <- nlsLM(ydata~exp_func(xdata, a1, b1, c1), start=list(a1=1, b1=0, c1=1))

data_plot <- ggplot(use_dat, aes(x = Year, y=Num.Sports)) + 
  geom_point() + 
  ylim(0, 40) + xlim(1948, 2020) +
  theme_minimal() +
  labs(title="Plot of the number of sports at the Olympic Game over the years \n starting from 1948", x="Year", y="Number of sports") + 
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1),
        axis.text.x = element_text(angle = 90, vjust = 0.5)) 

data_plot + geom_line(aes(x=xdata, y=predict(mod_lin)), color="red") + 
  geom_line(aes(x=xdata, y=predict(mod_quad)), color="orange") +
  geom_line(aes(x=xdata, y=predict(mod_cubic)), color="green") +
  geom_line(aes(x=xdata, y=predict(mod_exp)), color="blue")

model_list_1exp <- c("Linear", "Exponential")
model_list_123 <- c("Linear", "Quadratic", "Cubic")

anova_results1exp <- anova(mod_lin, mod_exp)
anova_results123 <- anova(mod_lin, mod_quad, mod_cubic) 

anova_res1exp <- cbind("Models"=model_list_1exp, anova_results1exp)
anova_res123 <- cbind("Models"=model_list_123, anova_results123)

anova_res1exp %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

anova_res123 %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

# Sports

## Top 10 Sports

```{r}
top10_sports <- data.frame(read.csv("top_10_sports.csv"))
head(top10_sports)
str(top10_sports)

ggplot(top10_sports, aes(x=Year, y=Mean_Weight, color=Sport)) + 
  geom_line() + 
  geom_point() +
  theme_minimal() +
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) + 
  labs(title="Plot of Olympic Weights [kg] Over The Years", x="year", y="weight [kg]") 

ggplot(top10_sports, aes(x=Year, y=Mean_Height, color=Sport)) + 
  geom_line() + 
  geom_point() +
  theme_minimal() +
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) + 
  labs(title="Plot of Olympic Heights [cm] Over The Years", x="year", y="height [kg]")
```

## Medalists

```{r}
loadPkg(tidyverse)

medalist <- olympic_data %>% 
  filter(!is.na(Medal)) 

medalist %>% 
  group_by(Medal) %>% 
  summarise(
    mean = mean(Age, na.rm = TRUE)
  ) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover")) 

summer_medalist_sex <-
  medalist %>%
  filter(Season == "Summer") %>% 
  group_by(Year, Sex) %>% 
    summarise(
    mean.Age = mean(Age, na.rm = TRUE)
  )

head(summer_medalist_sex)

summer_ages_gender <- ggplot(data = summer_medalist_sex, 
       mapping = aes(x = Year, y = mean.Age, group = Sex, color = Sex)) +
  geom_point(size=2) +
  geom_line() +
  labs(title="Average Age of Olympic Medalists \nin the Summer Games Separated by Gender",
       x="year", y="mean age [years]") +
  theme_minimal() +
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1))
summer_ages_gender
```

```{r}
filter_data <- function(df, sport) {
  df %>%
    filter(Sport == sport) %>% 
    group_by(Year, Sex) %>%
    summarise(
      mean.Age = mean(Age, na.rm = TRUE),
      sd.Age = sd(Age, na.rm = TRUE),
      mean.Weight = mean(Weight, na.rm = TRUE),
      sd.Weight = sd(Weight, na.rm = TRUE),
      mean.Height = mean(Height, na.rm = TRUE),
      sd.Height = sd(Height, na.rm = TRUE)
  )
}
```

```{r}
make_plots <- function(df, sport, varNameX, varNameY, xVal, yVal, unitX, unitY) {
  var_name_x <- eval(substitute(xVal),eval(df))
  var_name_y <- eval(substitute(yVal),eval(df))
  ggplot(df, mapping = aes(x = var_name_x, y = var_name_y, group = Sex, 
                                        color = Sex)) + geom_point(size=2) + 
    geom_line() +
    labs(title=paste(varNameY, " of Olympic Athletes in ", sport, "\nSeparated by Gender Over ", varNameX),
         x=paste(varNameX, unitX), y=paste(varNameY, unitY)) +
    theme_minimal() +
    theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) 
}
```

## Swimming

```{r}
sports <- olympic_data %>% 
  filter(!is.na(Sport)) 

swimming <- filter_data(sports, "Swimming")

make_plots(swimming, "Swimming", "Years", "Average Age", Year, mean.Age, "[years]","[years]")

make_plots(swimming, "Swimming", "Years", "Average Weight", Year, mean.Weight, "[years]","[kg]")

make_plots(swimming, "Swimming", "Years", "Average Height", Year, mean.Height, "[years]","[cm]")
```

### Female Athletes

```{r, results='markup'}
make_model_plots <- function(df, sport, varNameX, varNameY, xVal, yVal, unitX, unitY) {
  var_name_x <- eval(substitute(xVal), eval(df))
  var_name_y <- eval(substitute(yVal), eval(df))
  
  mod_lin <- nlsLM(var_name_y ~ lin_func(var_name_x, a1, b1), 
                   start=list(a1=1, b1=10))
  mod_quad <- nlsLM(var_name_y ~ quad_func(var_name_x, a1, b1, c1), 
                    start=list(a1=1, b1=1, c1=1))
  mod_cubic <- nlsLM(var_name_y ~ cubic_func(var_name_x, a1, b1, c1, d1), 
                     start=list(a1=1, b1=1, c1=1, d1=1))
  mod_exp <- nlsLM(var_name_y ~ exp_func(var_name_x, a1, b1, c1), 
                   start=list(a1=1, b1=0, c1=1))
  
  data_plot <- ggplot(df, aes(x = var_name_x, y = var_name_y, color = Sex)) + 
    geom_point() +
    theme_minimal() +
    labs(title=paste(varNameY, "of Olympic Athletes in", sport, "Over", varNameX), 
         x=paste(varNameX, unitX), y=paste(varNameY, unitY)) + 
    theme(panel.border = element_rect(colour = "black", fill=NA, size=1),
        axis.text.x = element_text(angle = 90, vjust = 0.5)) 
  
  plot_models <- data_plot + geom_line(aes(x=var_name_x, y=predict(mod_lin)), color="red") + 
    geom_line(aes(x=var_name_x, y=predict(mod_quad)), color="orange") +
    geom_line(aes(x=var_name_x, y=predict(mod_cubic)), color="green") +
    geom_line(aes(x=var_name_x, y=predict(mod_exp)), color="blue")
  
  anova_res1exp <- anova(mod_lin, mod_exp) 
  anova_res123 <- anova(mod_lin, mod_quad, mod_cubic) 
  
  return(list(plot_models, anova_res1exp, anova_res123))
}
```

```{r, results='markup'}
combine_anova <- function(anova_results, model_list) {
  anova_res <- cbind("Models"=model_list, anova_results)
  return(anova_res)
}
```


```{r, results='markup'}
# model_list_1exp <- c("Linear", "Exponential")
# model_list_123 <- c("Linear", "Quadratic", "Cubic")
# model_list <- c("Linear", "Exponential", "Quadratic", "Cubic")

index_F <- which(swimming$Sex == "F")
swimming_F <- swimming[index_F, ]

res_Swimming_Age <- make_model_plots(swimming_F, "Swimming", "Years", "Average Age", Year, mean.Age, "[years]","[years]")
res_Swimming_Age[[1]]

combine_anova(res_Swimming_Age[[2]], model_list_1exp) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

combine_anova(res_Swimming_Age[[3]], model_list_123) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

res_Swimming_Weight <- make_model_plots(swimming_F, "Swimming", "Years", "Average Weight", Year, mean.Weight, "[years]","[kg]")
res_Swimming_Weight[[1]]

combine_anova(res_Swimming_Weight[[2]], model_list_1exp) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

combine_anova(res_Swimming_Weight[[3]], model_list_123) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

res_Swimming_Height <- make_model_plots(swimming_F, "Swimming", "Years", "Average Height", Year, mean.Height, "[years]","[cm]")
res_Swimming_Height[[1]]

combine_anova(res_Swimming_Height[[2]], model_list_1exp) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

combine_anova(res_Swimming_Height[[3]], model_list_123) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

### Male Athletes

```{r, results='markup'}
index_M <- which(swimming$Sex == "M")
swimming_M <- swimming[index_M, ]

res_Swimming_Age <- make_model_plots(swimming_M, "Swimming", "Years", "Average Age", Year, mean.Age, "[years]","[years]")
res_Swimming_Age[[1]]

combine_anova(res_Swimming_Age[[2]], model_list_1exp) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

combine_anova(res_Swimming_Age[[3]], model_list_123) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

res_Swimming_Weight <- make_model_plots(swimming_M, "Swimming", "Years", "Average Weight", Year, mean.Weight, "[years]","[kg]")
res_Swimming_Weight[[1]]

combine_anova(res_Swimming_Weight[[2]], model_list_1exp) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

combine_anova(res_Swimming_Weight[[3]], model_list_123) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

res_Swimming_Height <- make_model_plots(swimming_M, "Swimming", "Years", "Average Height", Year, mean.Height, "[years]","[cm]")
res_Swimming_Height[[1]]

combine_anova(res_Swimming_Height[[2]], model_list_1exp) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

combine_anova(res_Swimming_Height[[3]], model_list_123) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

## Athletics

```{r}
athletics <- filter_data(sports, "Athletics")

make_plots(athletics, "Athletics", "Years", "Average Age", Year, mean.Age, "[years]","[years]")

make_plots(athletics, "Athletics", "Years", "Average Weight", Year, mean.Weight, "[years]","[kg]")

make_plots(athletics, "Athletics", "Years", "Average Height", Year, mean.Height, "[years]","[cm]")
```

### Female Athletes

```{r}
index_F <- which(athletics$Sex == "F")
athletics_F <- athletics[index_F, ]

res_Age <- make_model_plots(athletics_F, "Athletics", "Years", "Average Age", Year, mean.Age, "[years]","[years]")
res_Age[[1]]

combine_anova(res_Age[[2]], model_list_1exp) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

combine_anova(res_Age[[3]], model_list_123) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

res_Weight <- make_model_plots(athletics_F, "Athletics", "Years", "Average Weight", Year, mean.Weight, "[years]","[kg]")
res_Weight[[1]]

combine_anova(res_Weight[[2]], model_list_1exp) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

combine_anova(res_Weight[[3]], model_list_123) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

res_Height <- make_model_plots(athletics_F, "Athletics", "Years", "Average Height", Year, mean.Height, "[years]","[cm]")
res_Height[[1]]

combine_anova(res_Height[[2]], model_list_1exp) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

combine_anova(res_Height[[3]], model_list_123) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

### Male Athletes 

```{r}
index <- which(athletics$Sex == "M")
athletics_G <- athletics[index, ]

res_Age <- make_model_plots(athletics_G, "Athletics", "Years", "Average Age", Year, mean.Age, "[years]","[years]")
res_Age[[1]]

combine_anova(res_Age[[2]], model_list_1exp) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

combine_anova(res_Age[[3]], model_list_123) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

res_Weight <- make_model_plots(athletics_G, "Athletics", "Years", "Average Weight", Year, mean.Weight, "[years]","[kg]")
res_Weight[[1]]

combine_anova(res_Weight[[2]], model_list_1exp) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

combine_anova(res_Weight[[3]], model_list_123) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

res_Height <- make_model_plots(athletics_G, "Athletics", "Years", "Average Height", Year, mean.Height, "[years]","[cm]")
res_Height[[1]]

combine_anova(res_Height[[2]], model_list_1exp) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

combine_anova(res_Height[[3]], model_list_123) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

## Gymnastics

```{r}
gymnastics <- filter_data(sports, "Gymnastics")

make_plots(gymnastics, "Gymnastics", "Years", "Average Age", Year, mean.Age, "[years]","[years]")

make_plots(gymnastics, "Gymnastics", "Years", "Average Weight", Year, mean.Weight, "[years]","[kg]")

make_plots(gymnastics, "Gymnastics", "Years", "Average Height", Year, mean.Height, "[years]","[cm]")
```

### Female Athletes

```{r}
index <- which(gymnastics$Sex == "F")
gymnastics_G <- gymnastics[index, ]

res_Age <- make_model_plots(gymnastics_G, "Gymnastics", "Years", "Average Age", Year, mean.Age, "[years]","[years]")
res_Age[[1]]

combine_anova(res_Age[[2]], model_list_1exp) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

combine_anova(res_Age[[3]], model_list_123) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# res_Weight <- make_model_plots(gymnastics_G, "Gymnastics", "Years", "Average Weight", Year, mean.Weight, "[years]","[kg]")
# res_Weight[[1]]
# 
# combine_anova(res_Weight[[2]], model_list_1exp) %>%
#   kable() %>%
#   kable_styling(bootstrap_options = c("striped", "hover"))
# 
# combine_anova(res_Weight[[3]], model_list_123) %>%
#   kable() %>%
#   kable_styling(bootstrap_options = c("striped", "hover"))

# res_Height <- make_model_plots(gymnastics_G, "Gymnastics", "Years", "Average Height", Year, mean.Height, "[years]","[cm]")
# res_Height[[1]]
# 
# combine_anova(res_Height[[2]], model_list_1exp) %>%
#   kable() %>%
#   kable_styling(bootstrap_options = c("striped", "hover"))
# 
# combine_anova(res_Height[[3]], model_list_123) %>%
#   kable() %>%
#   kable_styling(bootstrap_options = c("striped", "hover"))
```

### Male Athletes

```{r}
index <- which(gymnastics$Sex == "M")
gymnastics_G <- gymnastics[index, ]

res_Age <- make_model_plots(gymnastics_G, "Gymnastics", "Years", "Average Age", Year, mean.Age, "[years]","[years]")
res_Age[[1]]

combine_anova(res_Age[[2]], model_list_1exp) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

combine_anova(res_Age[[3]], model_list_123) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# res_Weight <- make_model_plots(gymnastics_G, "Gymnastics", "Years", "Average Weight", Year, mean.Weight, "[years]","[kg]")
# res_Weight[[1]]
# 
# combine_anova(res_Weight[[2]], model_list_1exp) %>%
#   kable() %>%
#   kable_styling(bootstrap_options = c("striped", "hover"))
# 
# combine_anova(res_Weight[[3]], model_list_123) %>%
#   kable() %>%
#   kable_styling(bootstrap_options = c("striped", "hover"))

# res_Height <- make_model_plots(gymnastics_G, "Gymnastics", "Years", "Average Height", Year, mean.Height, "[years]","[cm]")
# res_Height[[1]]
# 
# combine_anova(res_Height[[2]], model_list_1exp) %>%
#   kable() %>%
#   kable_styling(bootstrap_options = c("striped", "hover"))
# 
# combine_anova(res_Height[[3]], model_list_123) %>%
#   kable() %>%
#   kable_styling(bootstrap_options = c("striped", "hover"))
```

## Rowing

```{r}
rowing <- filter_data(sports, "Rowing")

make_plots(rowing, "Rowing", "Years", "Average Age", Year, mean.Age, "[years]","[years]")

make_plots(rowing, "Rowing", "Years", "Average Weight", Year, mean.Weight, "[years]","[kg]")

make_plots(rowing, "Rowing", "Years", "Average Height", Year, mean.Height, "[years]","[cm]")
```

### Female Athletes

```{r}
index <- which(rowing$Sex == "F")
rowing_G <- rowing[index, ]

res_Age <- make_model_plots(rowing_G, "Rowing", "Years", "Average Age", Year, mean.Age, "[years]","[years]")
res_Age[[1]]

combine_anova(res_Age[[2]], model_list_1exp) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

combine_anova(res_Age[[3]], model_list_123) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

res_Weight <- make_model_plots(rowing_G, "Rowing", "Years", "Average Weight", Year, mean.Weight, "[years]","[kg]")
res_Weight[[1]]

combine_anova(res_Weight[[2]], model_list_1exp) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

combine_anova(res_Weight[[3]], model_list_123) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

res_Height <- make_model_plots(rowing_G, "Rowing", "Years", "Average Height", Year, mean.Height, "[years]","[cm]")
res_Height[[1]]

combine_anova(res_Height[[2]], model_list_1exp) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

combine_anova(res_Height[[3]], model_list_123) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

### Male Athletes

```{r}
index <- which(rowing$Sex == "M")
rowing_G <- rowing[index, ]

res_Age <- make_model_plots(rowing_G, "Rowing", "Years", "Average Age", Year, mean.Age, "[years]","[years]")
res_Age[[1]]

combine_anova(res_Age[[2]], model_list_1exp) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

combine_anova(res_Age[[3]], model_list_123) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

res_Weight <- make_model_plots(rowing_G, "Rowing", "Years", "Average Weight", Year, mean.Weight, "[years]","[kg]")
res_Weight[[1]]

combine_anova(res_Weight[[2]], model_list_1exp) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

combine_anova(res_Weight[[3]], model_list_123) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

res_Height <- make_model_plots(rowing_G, "Rowing", "Years", "Average Height", Year, mean.Height, "[years]","[cm]")
res_Height[[1]]

combine_anova(res_Height[[2]], model_list_1exp) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

combine_anova(res_Height[[3]], model_list_123) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

## Basketball

```{r}
basketball <- filter_data(sports, "Basketball")

make_plots(basketball, "Basketball", "Years", "Average Age", Year, mean.Age, "[years]","[years]")

make_plots(basketball, "Basketball", "Years", "Average Weight", Year, mean.Weight, "[years]","[kg]")

make_plots(basketball, "Basketball", "Years", "Average Height", Year, mean.Height, "[years]","[cm]")
```

### Female Athletes

```{r}
index <- which(basketball$Sex == "F")
basketball_G <- basketball[index, ]

res_Age <- make_model_plots(basketball_G, "Basketball", "Years", "Average Age", Year, mean.Age, "[years]","[years]")
res_Age[[1]]

combine_anova(res_Age[[2]], model_list_1exp) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

combine_anova(res_Age[[3]], model_list_123) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

res_Weight <- make_model_plots(basketball_G, "Basketball", "Years", "Average Weight", Year, mean.Weight, "[years]","[kg]")
res_Weight[[1]]

combine_anova(res_Weight[[2]], model_list_1exp) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

combine_anova(res_Weight[[3]], model_list_123) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

res_Height <- make_model_plots(basketball_G, "Basketball", "Years", "Average Height", Year, mean.Height, "[years]","[cm]")
res_Height[[1]]

combine_anova(res_Height[[2]], model_list_1exp) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

combine_anova(res_Height[[3]], model_list_123) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

### Male Athletes

```{r}
index <- which(basketball$Sex == "M")
basketball_G <- basketball[index, ]

res_Age <- make_model_plots(basketball_G, "Basketball", "Years", "Average Age", Year, mean.Age, "[years]","[years]")
res_Age[[1]]

combine_anova(res_Age[[2]], model_list_1exp) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

combine_anova(res_Age[[3]], model_list_123) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

res_Weight <- make_model_plots(basketball_G, "Basketball", "Years", "Average Weight", Year, mean.Weight, "[years]","[kg]")
res_Weight[[1]]

combine_anova(res_Weight[[2]], model_list_1exp) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

combine_anova(res_Weight[[3]], model_list_123) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

res_Height <- make_model_plots(basketball_G, "Basketball", "Years", "Average Height", Year, mean.Height, "[years]","[cm]")
res_Height[[1]]

combine_anova(res_Height[[2]], model_list_1exp) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

combine_anova(res_Height[[3]], model_list_123) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

## Fencing

```{r}
fencing <- filter_data(sports, "Fencing")

make_plots(fencing, "Fencing", "Years", "Average Age", Year, mean.Age, "[years]","[years]")

make_plots(fencing, "Fencing", "Years", "Average Weight", Year, mean.Weight, "[years]","[kg]")

make_plots(fencing, "Fencing", "Years", "Average Height", Year, mean.Height, "[years]","[cm]")
```

### Female Athletes

```{r}
index <- which(fencing$Sex == "F")
fencing_G <- fencing[index, ]

res_Age <- make_model_plots(fencing_G, "Fencing", "Years", "Average Age", Year, mean.Age, "[years]","[years]")
res_Age[[1]]

combine_anova(res_Age[[2]], model_list_1exp) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

combine_anova(res_Age[[3]], model_list_123) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

res_Weight <- make_model_plots(fencing_G, "Fencing", "Years", "Average Weight", Year, mean.Weight, "[years]","[kg]")
res_Weight[[1]]

combine_anova(res_Weight[[2]], model_list_1exp) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

combine_anova(res_Weight[[3]], model_list_123) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

res_Height <- make_model_plots(fencing_G, "Fencing", "Years", "Average Height", Year, mean.Height, "[years]","[cm]")
res_Height[[1]]

combine_anova(res_Height[[2]], model_list_1exp) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

combine_anova(res_Height[[3]], model_list_123) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

### Male Athletes

```{r}
index <- which(fencing$Sex == "M")
fencing_G <- fencing[index, ]

res_Age <- make_model_plots(fencing_G, "Fencing", "Years", "Average Age", Year, mean.Age, "[years]","[years]")
res_Age[[1]]

combine_anova(res_Age[[2]], model_list_1exp) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

combine_anova(res_Age[[3]], model_list_123) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

res_Weight <- make_model_plots(fencing_G, "Fencing", "Years", "Average Weight", Year, mean.Weight, "[years]","[kg]")
res_Weight[[1]]

combine_anova(res_Weight[[2]], model_list_1exp) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

combine_anova(res_Weight[[3]], model_list_123) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

res_Height <- make_model_plots(fencing_G, "Fencing", "Years", "Average Height", Year, mean.Height, "[years]","[cm]")
res_Height[[1]]

combine_anova(res_Height[[2]], model_list_1exp) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

combine_anova(res_Height[[3]], model_list_123) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

# References



```{r}
unloadPkg(tidyverse)
unloadPkg(kableExtra)
unloadPkg(minpack.lm)
unloadPkg(ggplot2)
```



