---
title: "Parameter Values for Models"
author: "Izzy Illari"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  rmdformats::readthedown:
    css: styles.css
    highlight: kate
    code_folding: hide
    number_sections: yes
    keep_tex: yes
    includes:
        after_body: github.html
bibliography: bibliography.bib
csl: american-institute-of-physics.csl
---

```{r clear, include=F}
rm(list = ls(all.names = TRUE))
gc() 
```

```{r basic, include=F}
# use this function to conveniently load libraries and work smoothly with knitting
# can add quietly=T option to the require() function
loadPkg = function(pkg, character.only = FALSE) { 
  if (!character.only) { pkg <- as.character(substitute(pkg)) }
  pkg <- ifelse(!character.only, as.character(substitute(pkg)) , pkg)  
  if (!require(pkg,character.only=T, quietly =T)) {  install.packages(substitute(pkg),dep=T); if(!require(pkg,character.only=T)) stop("Package not found") } 
}
loadPkg(knitr)

# unload/detact package when done using it
unloadPkg = function(pkg, character.only = FALSE) { 
  if(!character.only) { pkg <- as.character(substitute(pkg)) } 
  search_item <- paste("package", pkg,sep = ":") 
  while(search_item %in% search()) { detach(search_item, unload = TRUE, character.only = TRUE) } 
}
```

```{r setup, echo=FALSE, cache=FALSE}
loadPkg(knitr)
loadPkg(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

# Import the data

```{r import_data}
olympic_data <- data.frame(read.csv("olympic_data.csv"))
olympic_data$BMI.Category <- as.factor(olympic_data$BMI.Category)
olympic_data$Medal.No.Yes <- as.factor(olympic_data$Medal.No.Yes)
head(olympic_data)
str(olympic_data)
# loadPkg(psych)
# pairs.panels(olympic_data, #data frame 
#              method = "pearson", # correlation method
#              hist.col = "#00AFBB", # set histogram color, can use "#22AFBB", "red",
#              density = TRUE,  # show density plots
#              ellipses = TRUE # show correlation ellipses
#              )
# unloadPkg(psych)
```

```{r}
loadPkg(minpack.lm)
loadPkg(kableExtra)
loadPkg(ggplot2)
loadPkg(tidyverse)
```

Number of sports per year at the https://www.topendsports.com/events/summer/sports/number.htm [@WoodOlympEvents]

```{r}
sports_dat <- data.frame(read.csv("sports_per_year.csv"))

ggplot(sports_dat, aes(x = Year, y=Num.Sports)) + 
  geom_point() + 
  geom_line() + 
  ylim(0, 40) + 
  theme_minimal() +
  labs(title="Plot of the number of sports at the Olympic Game over the years", 
       x="Year", y="Number of sports") + 
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1),
        axis.text.x = element_text(angle = 90, vjust = 0.5)) 
```

There is clearly an upward trend, but no seasonal pattern. The data is also a little choppy at the beginning. Part of the explanation is that the data points are not evenly spaced. Most Olympic games are 4 years apart, but a few of them are just 2 years apart, and during World War I and World War II there were 8-year and 12-year gaps, respectively. Since time series data should be evenly spaced over time, we'll only look at data from 1948 on, when the Olympics started being held every 4 years without any interruptions.

```{r}
index_1948 <- which(sports_dat$Year == 1948)
use_dat <- sports_dat[12:nrow(sports_dat),]
ggplot(use_dat, aes(x = Year, y=Num.Sports)) + 
  geom_point() + 
  geom_line() + 
  ylim(0, 40) + 
  theme_minimal() +
  labs(title="Plot of the number of sports at the Olympic Game over the years \n starting from 1948", 
       x="Year", y="Number of sports") + 
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1),
        axis.text.x = element_text(angle = 90, vjust = 0.5)) 
```

# Creating models

I'm going to try 4 different models.

$$
y_{\text{linear}}(x) = ax+b \\
y_{\text{quadratic}}(x) = ax^2 + bx + c \\
y_{\text{cubic}}(x) = ax^3 + bx^2 + cx + d \\
y_{\text{exponential}}(x) = a\exp(bx) + c
$$

And I'll be able to use `ANOVA` to test the nested models: linear vs quadratic, and exponential growth vs s-curve (sigmoid). 

```{r func1}
# linear model function, it takes 3 inputs
lin_func <- function(x, a, b) {
  a*x+b
}
# quadratic model function, it takes 4 inputs
quad_func <- function(x, a, b, c) {
  a*x*x + b*x + c
}
# cubic model function, it takes 5 inputs
cubic_func <- function(x, a, b, c, d) {
  a*x*x*x + b*x*x + c*x + d 
}
# exponential model function, it takes 4 inputs
exp_func <- function(x, a, b, c) {
  a*exp(b*x) + c
}
```

## Chi-square "rule of thumb"

We consider $m$ equations that relate the $n$ random variables with values

$$
y_j = f(a_1,...,a_m,x_j) + \epsilon_j, ~~~~~ j = 1,...,n
$$

If we assume that only the $n-m$ random variables can fluctuate independently, and that the data uncertainties follow a Normal distribution, then the resulting chi-square is expected to be distributed according to

$$
\chi^2_{\nu}, ~~~~~ \nu = n - m
$$

Which we can compare to the "rule of thumb", which is

$$
\text{if } \frac{\chi^2}{n-m} \approx 1 ~~~ \text{where }\nu = n - m \implies \text{ "a good fit"}
$$

If we have $\nu$ independent RVs and the $x_i$ are each normally distributed with mean $\mu_i$ and variance $\sigma1_i^2$, then the chi-square is,

$$
\frac{\chi^2}{n-m} = \frac{\sum_{i=1}^{\nu} (x_i-\mu_i)^2/\sigma_i^2}{n-m}
$$

where $n$ is the length of our y data and $m$ is the degree of the polynomial. Using the above equation, we can calculate the values for our $\chi^2$ "rule of thumb".

```{r func2}
#calculates the chisquare/dof statistic
calc_chisquare <- function (model, yDat, degree) {
  # degree of exp = 1
  # deg of lin = 2
  # deg of quad = 3
  # deg of cubic = 4
  chisquare <- (sum((predict(model)-yDat) ^ 2)/(sd(yDat) ^ 2))/(length(yDat)-degree)
  chisquare <- round(chisquare, 3)
  return(chisquare)
}
```

## Model Fits

### Number of events over time

Now I will try the model fits on the number of events per Olympic Games data. 

```{r}
xdata <- use_dat$Year
ydata <- use_dat$Num.Sports

mod_lin <- nlsLM(ydata~lin_func(xdata, a1, b1), start=list(a1=1, b1=10))
summary(mod_lin)
summary(lm(Num.Sports~Year, data=use_dat))
mod_quad <- nlsLM(ydata~quad_func(xdata, a1, b1, c1), start=list(a1=1, b1=1, c1=1))
mod_cubic <- nlsLM(ydata~cubic_func(xdata, a1, b1, c1, d1), start=list(a1=1, b1=1, c1=1, d1=1))
mod_exp <- nlsLM(ydata~exp_func(xdata, a1, b1, c1), start=list(a1=1, b1=0, c1=1))

data_plot <- ggplot(use_dat, aes(x = Year, y=Num.Sports)) + 
  geom_point() + 
  ylim(0, 40) + xlim(1948, 2020) +
  theme_minimal() +
  labs(title="Plot of the number of sports at the Olympic Game over the years \n starting from 1948", x="Year", y="Number of sports") + 
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1),
        axis.text.x = element_text(angle = 90, vjust = 0.5)) 

data_plot + geom_line(aes(x=xdata, y=predict(mod_lin)), color="red") + 
  geom_line(aes(x=xdata, y=predict(mod_quad)), color="orange") +
  geom_line(aes(x=xdata, y=predict(mod_cubic)), color="green") +
  geom_line(aes(x=xdata, y=predict(mod_exp)), color="blue")
```

```{r}
model_list_1exp <- c("Linear", "Exponential")
model_list_123 <- c("Linear", "Quadratic", "Cubic")

anova_results1exp <- anova(mod_lin, mod_exp)
anova_results123 <- anova(mod_lin, mod_quad, mod_cubic) 

anova_res1exp <- cbind("Models"=model_list_1exp, anova_results1exp)
anova_res123 <- cbind("Models"=model_list_123, anova_results123)

anova_res1exp %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

anova_res123 %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

```{r, echo=F}
loadPkg(tidyverse)
#loadPkg(ggplot2)
loadPkg(stringr)
#library(infer)
loadPkg(dplyr)
loadPkg(infer)
```

```{r}
use_dat$Num.Sports <- as.factor(use_dat$Num.Sports)
obs_chisq <- use_dat %>%
  specify(Num.Sports ~ Year) %>% # alt: response = origin, explanatory = season
  calculate(stat = "Chisq")
#obs_chisq <- use_dat %>% 
#  chisq_stat(formula = Num.Sports ~ Year)
obs_chisq
```

# Swiming

```{r filterDataFunc}
#function to filter data and creates a new data frame
#input is the data frame and the name of the sport
#eg use the following: filter_data(olympic_data, "Rowing")
filter_data <- function(df, sport) {
  #filters input dataframe by the Sport and groups the results by Year and Sex
  df %>% 
    filter(Sport == sport) %>% 
    group_by(Year, Sex) %>%
    summarise(
      #calculates mean age for each year
      mean.Age = mean(Age, na.rm = TRUE), 
      #calculates standard deviation of age for each year
      sd.Age = sd(Age, na.rm = TRUE),
      #calculates mean weight for each year
      mean.Weight = mean(Weight, na.rm = TRUE),
      #calculates standard deviation of weight for each year
      sd.Weight = sd(Weight, na.rm = TRUE),
      #calculates mean height for each year
      mean.Height = mean(Height, na.rm = TRUE),
      #calculates standard deviation of height for each year
      sd.Height = sd(Height, na.rm = TRUE)
  )
}
```

```{r makePlotsFunc}
#function to make the ggplots using the data frames created by the above function
#input should be in the following style:
# make_plots(swimming, "Swimming", "Years", "Average Weight", Year, mean.Weight, "[years]","[kg]")
make_plots <- function(df, sport, varNameX, varNameY, xVal, yVal, unitX, unitY) {
  var_name_x <- eval(substitute(xVal),eval(df))
  var_name_y <- eval(substitute(yVal),eval(df))
  ggplot(df, mapping = aes(x = var_name_x, y = var_name_y, group = Sex, 
                                        color = Sex)) + geom_point(size=2) + 
    geom_line() +
    labs(title=paste(varNameY, " of Olympic Athletes in ", sport, "\nSeparated by Gender Over ", varNameX),
         x=paste(varNameX, unitX), y=paste(varNameY, unitY)) +
    theme_minimal() +
    theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) 
}
```

```{r}
sports <- olympic_data %>% 
  filter(!is.na(Sport)) 

swimming <- filter_data(sports, "Swimming")

make_plots(swimming, "Swimming", "Years", "Average Age", Year, mean.Age, "[years]","[years]")

make_plots(swimming, "Swimming", "Years", "Average Weight", Year, mean.Weight, "[years]","[kg]")

make_plots(swimming, "Swimming", "Years", "Average Height", Year, mean.Height, "[years]","[cm]")
```

```{r modelFits}
# linear model function, it takes 3 inputs
lin_func <- function(x, a, b) {
  a*x+b
}
# quadratic model function, it takes 4 inputs
quad_func <- function(x, a, b, c) {
  a*x*x + b*x + c
}
# cubic model function, it takes 5 inputs
cubic_func <- function(x, a, b, c, d) {
  a*x*x*x + b*x*x + c*x + d 
}
# exponential model function, it takes 4 inputs
exp_func <- function(x, a, b, c) {
  a*exp(b*x) + c
}
```

## Female Swimmers

```{r plotModels, results='markup'}
#make plots with the model fits
#to use the function write something like the following:
#make_model_plots(swimming_F, "Swimming", "Years", "Average Age", Year, mean.Age, "[years]","[years]")
make_model_plots <- function(df, sport, varNameX, varNameY, xVal, yVal, unitX, unitY) {
  var_name_x <- eval(substitute(xVal), eval(df))
  var_name_y <- eval(substitute(yVal), eval(df))
  #linear model fit
  mod_lin <- nlsLM(var_name_y ~ lin_func(var_name_x, a1, b1), 
                   start=list(a1=1, b1=10))
  #quadratic model fit
  mod_quad <- nlsLM(var_name_y ~ quad_func(var_name_x, a1, b1, c1), 
                    start=list(a1=1, b1=1, c1=1))
  #cubic model fit
  mod_cubic <- nlsLM(var_name_y ~ cubic_func(var_name_x, a1, b1, c1, d1), 
                     start=list(a1=1, b1=1, c1=1, d1=1))
  #exponential model fit
  mod_exp <- nlsLM(var_name_y ~ exp_func(var_name_x, a1, b1, c1), 
                   start=list(a1=1, b1=0, c1=1))
  #plot data without models
  data_plot <- ggplot(df, aes(x = var_name_x, y = var_name_y, color = Sex)) + 
    geom_point() +
    theme_minimal() +
    labs(title=paste(varNameY, "of Olympic Athletes in", sport, "Over", varNameX), 
         x=paste(varNameX, unitX), y=paste(varNameY, unitY)) + 
    theme(panel.border = element_rect(colour = "black", fill=NA, size=1),
        axis.text.x = element_text(angle = 90, vjust = 0.5)) 
  #manually add each line for a model
  plot_models <- data_plot + geom_line(aes(x=var_name_x, y=predict(mod_lin)), color="red") + 
    geom_line(aes(x=var_name_x, y=predict(mod_quad)), color="orange") +
    geom_line(aes(x=var_name_x, y=predict(mod_cubic)), color="green") +
    geom_line(aes(x=var_name_x, y=predict(mod_exp)), color="blue")
  #calculate the anova results
  anova_res1exp <- anova(mod_lin, mod_exp) 
  anova_res123 <- anova(mod_lin, mod_quad, mod_cubic) 
  #return the plot with model lines, the anova results, and chisquare/dof statistic for each model
  return(list(plot_models, anova_res1exp, anova_res123, calc_chisquare(mod_exp, ydata, 1), calc_chisquare(mod_lin, ydata, 2), calc_chisquare(mod_quad, ydata, 3), calc_chisquare(mod_cubic, ydata, 4)))
}
#output will be in a nested list
#for example you would do the following to call the results:
# results <- make_model_plots(swimming_F, "Swimming", "Years", "Average Weight", Year, mean.Weight, "[years]","[kg]")
# and to call the plot results you would write
# plot <- results[[1]]
# anova_linear_exponential <- results[[2]]
# anova_lin_quad_cubic <- results[[3]]
# chisquare_exponential <- results[[4]]
# chisquare_linear <- results[[5]]
# chisquare_quadratic <- results[[6]]
# chisquare_cubic <- results[[7]]
```

```{r combineANOVAfunc, results='markup'}
#function to combine the anove results with the model list
#same idea as the chisquare function earlier
combine_anova <- function(anova_results, model_list) {
  anova_res <- cbind("Models"=model_list, anova_results)
  return(anova_res)
}
```

```{r}
index_F <- which(swimming$Sex == "F")
swimming_F <- swimming[index_F, ]

res_Swimming_Age <- make_model_plots(swimming_F, "Swimming", "Years", "Average Age", Year, mean.Age, "[years]","[years]")
res_Swimming_Age[[1]]

combine_anova(res_Swimming_Age[[2]], model_list_1exp) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

combine_anova(res_Swimming_Age[[3]], model_list_123) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```



# References
