---
title: "Olympic Data"
author: "Izzy Illari"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  rmdformats::readthedown:
    css: styles.css
    highlight: kate
    code_folding: hide
    number_sections: yes
    keep_tex: yes
    includes:
        after_body: github.html
bibliography: bibliography.bib
csl: american-institute-of-physics.csl
---

```{r clear, include=F}
rm(list = ls(all.names = TRUE))
gc() 
```


```{r basic, include=F}
# use this function to conveniently load libraries and work smoothly with knitting
# can add quietly=T option to the require() function
loadPkg = function(pkg, character.only = FALSE) { 
  if (!character.only) { pkg <- as.character(substitute(pkg)) }
  pkg <- ifelse(!character.only, as.character(substitute(pkg)) , pkg)  
  if (!require(pkg,character.only=T, quietly =T)) {  install.packages(substitute(pkg),dep=T); if(!require(pkg,character.only=T)) stop("Package not found") } 
}
loadPkg(knitr)

# unload/detact package when done using it
unloadPkg = function(pkg, character.only = FALSE) { 
  if(!character.only) { pkg <- as.character(substitute(pkg)) } 
  search_item <- paste("package", pkg,sep = ":") 
  while(search_item %in% search()) { detach(search_item, unload = TRUE, character.only = TRUE) } 
}
```


```{r setup, echo=FALSE, cache=FALSE}
loadPkg(knitr)
loadPkg(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

# Import the data

```{r import_data}
olympic_data <- data.frame(read.csv("olympic_data.csv"))
olympic_data$BMI.Category <- as.factor(olympic_data$BMI.Category)
olympic_data$Medal.No.Yes <- as.factor(olympic_data$Medal.No.Yes)
head(olympic_data)
str(olympic_data)
```

Done. 

# Time Series

## Number of Events 

Lets look to see

```{r}
data_year <- split(olympic_data, olympic_data$Year)

list_year <- list(1960, 1964, 1968, 1972, 1976, 1980, 1984, 1988, 1992, 1994, 1996, 1998, 2000, 2002, 2004, 2006, 2008, 2010, 2012, 2014, 2016)
index <- list()
list_df <- list()
list_count_s <- list()
list_count_e <- list()
combined_df <- list()
sum_count_s <- list()
sum_count_e <- list()
unique_s <- list()
unique_e <- list()
sum_unique_s <- list()
sum_unique_e <- list()

loadPkg(plyr)
for(i in 1:length(list_year)) {
  index[[i]] <- which(olympic_data$Year == list_year[i])
  list_df[[i]] <- olympic_data[index[[i]],]
  unique_s[[i]] <- unique(list_df[[i]]$Sport)
  sum_unique_s[i] <- length(unique_s[[i]])
  unique_e[[i]] <- unique(list_df[[i]]$Event)
  sum_unique_e[i] <- length(unique_e[[i]])
  list_count_s[[i]] <- count(list_df[[i]], 'Sport')
  sum_count_s[i] <- sum(list_count_s[[i]]$freq)
  list_count_e[[i]] <- count(list_df[[i]], 'Event')
  sum_count_e[i] <- sum(list_count_e[[i]]$freq)
}
count_se <- cbind(unlist(list_year), unlist(sum_count_s), unlist(sum_unique_s), unlist(sum_count_e), unlist(sum_unique_e))
count_se <- as.data.frame(count_se)
colnames(count_se) <- c("Year", "Num.Sports", "Tot.Sports", "Num.Events", "Tot.Events")

loadPkg(ggplot2)
ggplot(count_se, aes(x=Year, y=Num.Sports)) + geom_point() + geom_line()
ggplot(count_se, aes(x=Year, y=Num.Events)) + geom_point() + geom_line()
ggplot(count_se, aes(x=Year, y=Tot.Sports)) + geom_point() + geom_line()
ggplot(count_se, aes(x=Year, y=Tot.Events)) + geom_point() + geom_line()
```

Number of sports per year at the https://www.topendsports.com/events/summer/sports/number.htm


```{r}
sports_dat <- data.frame(read.csv("sports_per_year.csv"))

ggplot(sports_dat, aes(x = Year, y=Num.Sports)) + 
  geom_point() + 
  geom_line() + 
  ylim(0, 40) + 
  theme_minimal() +
  labs(title="Plot of the number of sports at the Olympic Game over the years", 
       x="Year", y="Number of sports") + 
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1),
        axis.text.x = element_text(angle = 90, vjust = 0.5)) 
```

There is clearly an upward trend, but no seasonal pattern. The data is also a little choppy at the beginning. Part of the explanation is that the data points are not evenly spaced. Most Olympic games are 4 years apart, but a few of them are just 2 years apart, and during World War I and World War II there were 8-year and 12-year gaps, respectively. Since time series data should be evenly spaced over time, we'll only look at data from 1948 on, when the Olympics started being held every 4 years without any interruptions.

```{r}
index_1948 <- which(sports_dat$Year == 1948)
use_dat <- sports_dat[12:nrow(sports_dat),]
ggplot(use_dat, aes(x = Year, y=Num.Sports)) + 
  geom_point() + 
  geom_line() + 
  ylim(0, 40) + 
  theme_minimal() +
  labs(title="Plot of the number of sports at the Olympic Game over the years \n starting from 1948", 
       x="Year", y="Number of sports") + 
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1),
        axis.text.x = element_text(angle = 90, vjust = 0.5)) 
```

Lets see if I can build a time series using our data.

```{r}
loadPkg(stats)
start_yr <- as.integer(use_dat[1,]$Year)
num_row <- as.integer(nrow(use_dat))
end_yr <- as.integer(use_dat[num_row,]$Year)
matrix <- use_dat[,1:2]
matrix <- data.matrix(matrix)
ts_olympics <- ts(matrix, start = start_yr, end = end_yr)
ts_olympics
plot(ts_olympics)
```

## Creating the models

I'm going to try 4 different models.

$$
y_{\text{linear}}(x) = ax+b \\
y_{\text{quadratic}}(x) = ax^2 + bx + c \\
y_{\text{exponential}}(x) = a\exp(bx) + c \\
y_{\text{cubic}}(x) = ax^3 + bx^2 + cx + d 
$$

And I'll be able to use `ANOVA` to test the nested models: linear vs quadratic, and exponential growth vs s-curve (sigmoid). 

```{r}
lin_func <- function(x, a, b) {
  a*x+b
}

quad_func <- function(x, a, b, c) {
  a*x*x + b*x + c
}

cubic_func <- function(x, a, b, c, d) {
  a*x*x*x + b*x*x + c*x + d 
}

exp_func <- function(x, a, b, c) {
  a*exp(b*x) + c
}
```

```{r}
loadPkg(minpack.lm)
xdata <- use_dat$Year
ydata <- use_dat$Num.Sports

mod_lin <- nlsLM(ydata~lin_func(xdata, a1, b1), start=list(a1=1, b1=10))
mod_quad <- nlsLM(ydata~quad_func(xdata, a1, b1, c1), start=list(a1=1, b1=1, c1=1))
mod_cubic <- nlsLM(ydata~cubic_func(xdata, a1, b1, c1, d1), start=list(a1=1, b1=1, c1=1, d1=1))
mod_exp <- nlsLM(ydata~exp_func(xdata, a1, b1, c1), start=list(a1=1, b1=0, c1=1))

data_plot <- ggplot(use_dat, aes(x = Year, y=Num.Sports)) + 
  geom_point() + 
  ylim(0, 40) + xlim(1948, 2020) +
  theme_minimal() +
  labs(title="Plot of the number of sports at the Olympic Game over the years \n starting from 1948", x="Year", y="Number of sports") + 
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1),
        axis.text.x = element_text(angle = 90, vjust = 0.5)) 

data_plot + geom_line(aes(x=xdata, y=predict(mod_lin)), color="red") + 
  geom_line(aes(x=xdata, y=predict(mod_quad)), color="orange") +
  geom_line(aes(x=xdata, y=predict(mod_cubic)), color="green") +
  geom_line(aes(x=xdata, y=predict(mod_exp)), color="blue")
```

These models all look fairly similar. Lets check using ANOVA.

```{r anova_events}
loadPkg(kableExtra)
anova(mod_lin, mod_quad, mod_exp, mod_cubic) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

Linear model preferred. Nothing gained from adding complexity to the model. 

Lets look at the top 10 sports by number of participants. 

```{r}
freq_sports <- count(olympic_data, "Sport")
freq_sports <- freq_sports[order(-freq_sports$freq),]
top_10_sports <- freq_sports[1:10,]
top_10_sports
top_10_sports %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover")) 
```


```{r}
ggplot(olympic_data, aes(x=Year, y=Weight, color=Sex)) + geom_point()
```

I need to subset the data because I keep getting the following error: "Error: vector memory exhausted (limit reached?)". I will drop the following variables: NOC, Decade, ID, First.Name, Name, BMI, BMI.Category, Games, City, Event. I will only focus on the top ten sports. 

```{r}
#head(olympic_data)
#colnames(olympic_data)
less_data <- subset(olympic_data, select = -c(NOC, Decade, ID, First.Name, Name, BMI, BMI.Category, Games, City, Event) )
nrow(less_data)

index_Athletics <- which(less_data$Sport == "Athletics")
index_Swimming <- which(less_data$Sport == "Swimming")
index_Gymnastics <- which(less_data$Sport == "Gymnastics")
index_Cross_Country_Skiing <- which(less_data$Sport == "Cross Country Skiing")
index_Alpine_Skiing <- which(less_data$Sport == "Alpine Skiing")
index_Cycling <- which(less_data$Sport == "Cycling")
index_Rowing <- which(less_data$Sport == "Rowing")
index_Shooting <- which(less_data$Sport == "Shooting")
index_Fencing <- which(less_data$Sport == "Fencing")
index_Canoeing <- which(less_data$Sport == "Canoeing")

index_combo_10 <- c(index_Athletics, index_Swimming, index_Gymnastics, index_Cross_Country_Skiing, index_Alpine_Skiing, index_Rowing, index_Shooting, index_Fencing, index_Canoeing)
index_combo_5 <- c(index_Athletics, index_Swimming, index_Gymnastics, index_Cross_Country_Skiing, index_Alpine_Skiing)
less_data <- less_data[index_combo_5,]
nrow(less_data)
#colnames(less_data)
```

```{r cor_packages, include=F}
loadPkg(reshape2)
```

```{r correlation_matrix, includ=F}
corr_plot <- function(dataName, methodUse) {
  corr <- round(cor(dataName, method = methodUse, use = "complete.obs"), 2)
  melted_cormat <- melt(corr)
  
  ggplot(data = melted_cormat, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile(color = "white") + 
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 0, 
                       limit = c(-1,1), space = "Lab", name=methodUse) + 
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 1, size = 5, hjust = 1), 
        axis.text.y = element_text(vjust = 1, size = 5, hjust = 1), 
        axis.title.x = element_blank(), axis.title.y = element_blank()) + 
  coord_fixed() + 
  geom_text(aes(Var2, Var1, label = value), color = "black", size = 2)
}
```

Lets try making logistic regression models for `Weight` and `Height`. 

```{r}
# logit_Weight <- glm(Weight ~ ., data = less_data, family = "binomial")
# summary(logit_Weight)
# logit_Height <- glm(Height ~ ., data = less_data, family = "binomial")
# summary(logit_Height)
```

```{r}
# data_sport <- split(less_data, less_data$Sport)
# # Athletics
# data_Athletics_gender <- split(data_sport$Athletics, data_sport$Athletics$Sex)
# data_Athletics_F_year <- split(data_Athletics_gender$F, data_Athletics_gender$F$Year)
# data_Athletics_M_year <- split(data_Athletics_gender$"M", data_Athletics_gender$"M"$Year)
# years_F_Athletics <- names(data_Athletics_F_year)
# years_M_Athletics <- names(data_Athletics_M_year)
# mean_years_weight_F_Athletics <- list()
# sd_years_weight_F_Athletics <- list()
# mean_years_height_F_Athletics <- list()
# sd_years_height_F_Athletics <- list()
# mean_years_weight_M_Athletics <- list()
# sd_years_weight_M_Athletics <- list()
# mean_years_height_M_Athletics <- list()
# sd_years_height_M_Athletics <- list()
# # women data
# for (i in 1:length(data_Athletics_F_year)) {
# 	mean_years_weight_F_Athletics[[i]] <- mean(data_Athletics_F_year[[i]]$Weight)
# 	sd_years_weight_F_Athletics[[i]] <- sd(data_Athletics_F_year[[i]]$Weight)
# 	mean_years_height_F_Athletics[[i]] <- mean(data_Athletics_F_year[[i]]$Height)
# 	sd_years_height_F_Athletics[[i]] <- sd(data_Athletics_F_year[[i]]$Height)
# }
# # men data
# for (i in 1:length(data_Athletics_M_year)) {
# 	mean_years_weight_M_Athletics[[i]] <- mean(data_Athletics_M_year[[i]]$Weight)
# 	sd_years_weight_M_Athletics[[i]] <- sd(data_Athletics_M_year[[i]]$Weight)
# 	mean_years_height_M_Athletics[[i]] <- mean(data_Athletics_M_year[[i]]$Height)
# 	sd_years_height_M_Athletics[[i]] <- sd(data_Athletics_M_year[[i]]$Height)
# }
# # make data frames
# Athletics_F_data <- do.call(rbind, Map(data.frame, Year = years_F_Athletics, Mean_Weight = mean_years_weight_F_Athletics, StdDev_Weight = sd_years_weight_F_Athletics, Mean_Height = mean_years_height_F_Athletics, StdDev_Height = sd_years_height_F_Athletics))
# Athletics_M_data <- do.call(rbind, Map(data.frame, Year = years_M_Athletics, Mean_Weight = mean_years_weight_F_Athletics, StdDev_Weight = sd_years_weight_F_Athletics, Mean_Height = mean_years_height_F_Athletics, StdDev_Height = sd_years_height_F_Athletics))
# # Swimming
# data_Swimming_gender <- split(data_sport$Swimming, data_sport$Swimming$Sex)
# data_Swimming_F_year <- split(data_Swimming_gender$F, data_Swimming_gender$F$Year)
# data_Swimming_M_year <- split(data_Swimming_gender$"M", data_Swimming_gender$"M"$Year)
# years_F_Swimming <- names(data_Swimming_F_year)
# years_M_Swimming <- names(data_Swimming_M_year)
# mean_years_weight_F_Swimming <- list()
# sd_years_weight_F_Swimming <- list()
# mean_years_height_F_Swimming <- list()
# sd_years_height_F_Swimming <- list()
# mean_years_weight_M_Swimming <- list()
# sd_years_weight_M_Swimming <- list()
# mean_years_height_M_Swimming <- list()
# sd_years_height_M_Swimming <- list()
# # women data
# for (i in 1:length(data_Swimming_F_year)) {
# 	mean_years_weight_F_Swimming[[i]] <- mean(data_Swimming_F_year[[i]]$Weight)
# 	sd_years_weight_F_Swimming[[i]] <- sd(data_Swimming_F_year[[i]]$Weight)
# 	mean_years_height_F_Swimming[[i]] <- mean(data_Swimming_F_year[[i]]$Height)
# 	sd_years_height_F_Swimming[[i]] <- sd(data_Swimming_F_year[[i]]$Height)
# }
# # men data
# for (i in 1:length(data_Swimming_M_year)) {
# 	mean_years_weight_M_Swimming[[i]] <- mean(data_Swimming_M_year[[i]]$Weight)
# 	sd_years_weight_M_Swimming[[i]] <- sd(data_Swimming_M_year[[i]]$Weight)
# 	mean_years_height_M_Swimming[[i]] <- mean(data_Swimming_M_year[[i]]$Height)
# 	sd_years_height_M_Swimming[[i]] <- sd(data_Swimming_M_year[[i]]$Height)
# }
# # make data frames
# Swimming_F_data <- do.call(rbind, Map(data.frame, Year = years_F_Swimming, Mean_Weight = mean_years_weight_F_Swimming, StdDev_Weight = sd_years_weight_F_Swimming, Mean_Height = mean_years_height_F_Swimming, StdDev_Height = sd_years_height_F_Swimming))
# Swimming_M_data <- do.call(rbind, Map(data.frame, Year = years_M_Swimming, Mean_Weight = mean_years_weight_F_Swimming, StdDev_Weight = sd_years_weight_F_Swimming, Mean_Height = mean_years_height_F_Swimming, StdDev_Height = sd_years_height_F_Swimming))
# # Gymnastics
# data_Gymnastics_gender <- split(data_sport$Gymnastics, data_sport$Gymnastics$Sex)
# data_Gymnastics_F_year <- split(data_Gymnastics_gender$F, data_Gymnastics_gender$F$Year)
# data_Gymnastics_M_year <- split(data_Gymnastics_gender$"M", data_Gymnastics_gender$"M"$Year)
# years_F_Gymnastics <- names(data_Gymnastics_F_year)
# years_M_Gymnastics <- names(data_Gymnastics_M_year)
# mean_years_weight_F_Gymnastics <- list()
# sd_years_weight_F_Gymnastics <- list()
# mean_years_height_F_Gymnastics <- list()
# sd_years_height_F_Gymnastics <- list()
# mean_years_weight_M_Gymnastics <- list()
# sd_years_weight_M_Gymnastics <- list()
# mean_years_height_M_Gymnastics <- list()
# sd_years_height_M_Gymnastics <- list()
# # women data
# for (i in 1:length(data_Gymnastics_F_year)) {
# 	mean_years_weight_F_Gymnastics[[i]] <- mean(data_Gymnastics_F_year[[i]]$Weight)
# 	sd_years_weight_F_Gymnastics[[i]] <- sd(data_Gymnastics_F_year[[i]]$Weight)
# 	mean_years_height_F_Gymnastics[[i]] <- mean(data_Gymnastics_F_year[[i]]$Height)
# 	sd_years_height_F_Gymnastics[[i]] <- sd(data_Gymnastics_F_year[[i]]$Height)
# }
# # men data
# for (i in 1:length(data_Gymnastics_M_year)) {
# 	mean_years_weight_M_Gymnastics[[i]] <- mean(data_Gymnastics_M_year[[i]]$Weight)
# 	sd_years_weight_M_Gymnastics[[i]] <- sd(data_Gymnastics_M_year[[i]]$Weight)
# 	mean_years_height_M_Gymnastics[[i]] <- mean(data_Gymnastics_M_year[[i]]$Height)
# 	sd_years_height_M_Gymnastics[[i]] <- sd(data_Gymnastics_M_year[[i]]$Height)
# }
# # make data frames
# Gymnastics_F_data <- do.call(rbind, Map(data.frame, Year = years_F_Gymnastics, Mean_Weight = mean_years_weight_F_Gymnastics, StdDev_Weight = sd_years_weight_F_Gymnastics, Mean_Height = mean_years_height_F_Gymnastics, StdDev_Height = sd_years_height_F_Gymnastics))
# Gymnastics_M_data <- do.call(rbind, Map(data.frame, Year = years_M_Gymnastics, Mean_Weight = mean_years_weight_F_Gymnastics, StdDev_Weight = sd_years_weight_F_Gymnastics, Mean_Height = mean_years_height_F_Gymnastics, StdDev_Height = sd_years_height_F_Gymnastics))
# # Cross.Country.Skiing
# data_Cross.Country.Skiing_gender <- split(data_sport$"Cross Country Skiing", data_sport$"Cross Country Skiing"$Sex)
# data_Cross.Country.Skiing_F_year <- split(data_Cross.Country.Skiing_gender$F, data_Cross.Country.Skiing_gender$F$Year)
# data_Cross.Country.Skiing_M_year <- split(data_Cross.Country.Skiing_gender$"M", data_Cross.Country.Skiing_gender$"M"$Year)
# years_F_Cross.Country.Skiing <- names(data_Cross.Country.Skiing_F_year)
# years_M_Cross.Country.Skiing <- names(data_Cross.Country.Skiing_M_year)
# mean_years_weight_F_Cross.Country.Skiing <- list()
# sd_years_weight_F_Cross.Country.Skiing <- list()
# mean_years_height_F_Cross.Country.Skiing <- list()
# sd_years_height_F_Cross.Country.Skiing <- list()
# mean_years_weight_M_Cross.Country.Skiing <- list()
# sd_years_weight_M_Cross.Country.Skiing <- list()
# mean_years_height_M_Cross.Country.Skiing <- list()
# sd_years_height_M_Cross.Country.Skiing <- list()
# # women data
# for (i in 1:length(data_Cross.Country.Skiing_F_year)) {
# 	mean_years_weight_F_Cross.Country.Skiing[[i]] <- mean(data_Cross.Country.Skiing_F_year[[i]]$Weight)
# 	sd_years_weight_F_Cross.Country.Skiing[[i]] <- sd(data_Cross.Country.Skiing_F_year[[i]]$Weight)
# 	mean_years_height_F_Cross.Country.Skiing[[i]] <- mean(data_Cross.Country.Skiing_F_year[[i]]$Height)
# 	sd_years_height_F_Cross.Country.Skiing[[i]] <- sd(data_Cross.Country.Skiing_F_year[[i]]$Height)
# }
# # men data
# for (i in 1:length(data_Cross.Country.Skiing_M_year)) {
# 	mean_years_weight_M_Cross.Country.Skiing[[i]] <- mean(data_Cross.Country.Skiing_M_year[[i]]$Weight)
# 	sd_years_weight_M_Cross.Country.Skiing[[i]] <- sd(data_Cross.Country.Skiing_M_year[[i]]$Weight)
# 	mean_years_height_M_Cross.Country.Skiing[[i]] <- mean(data_Cross.Country.Skiing_M_year[[i]]$Height)
# 	sd_years_height_M_Cross.Country.Skiing[[i]] <- sd(data_Cross.Country.Skiing_M_year[[i]]$Height)
# }
# # make data frames
# Cross.Country.Skiing_F_data <- do.call(rbind, Map(data.frame, Year = years_F_Cross.Country.Skiing, Mean_Weight = mean_years_weight_F_Cross.Country.Skiing, StdDev_Weight = sd_years_weight_F_Cross.Country.Skiing, Mean_Height = mean_years_height_F_Cross.Country.Skiing, StdDev_Height = sd_years_height_F_Cross.Country.Skiing))
# Cross.Country.Skiing_M_data <- do.call(rbind, Map(data.frame, Year = years_M_Cross.Country.Skiing, Mean_Weight = mean_years_weight_F_Cross.Country.Skiing, StdDev_Weight = sd_years_weight_F_Cross.Country.Skiing, Mean_Height = mean_years_height_F_Cross.Country.Skiing, StdDev_Height = sd_years_height_F_Cross.Country.Skiing))
# # Alpine.Skiing
# data_Alpine.Skiing_gender <- split(data_sport$"Alpine Skiing", data_sport$"Alpine Skiing"$Sex)
# data_Alpine.Skiing_F_year <- split(data_Alpine.Skiing_gender$F, data_Alpine.Skiing_gender$F$Year)
# data_Alpine.Skiing_M_year <- split(data_Alpine.Skiing_gender$"M", data_Alpine.Skiing_gender$"M"$Year)
# years_F_Alpine.Skiing <- names(data_Alpine.Skiing_F_year)
# years_M_Alpine.Skiing <- names(data_Alpine.Skiing_M_year)
# mean_years_weight_F_Alpine.Skiing <- list()
# sd_years_weight_F_Alpine.Skiing <- list()
# mean_years_height_F_Alpine.Skiing <- list()
# sd_years_height_F_Alpine.Skiing <- list()
# mean_years_weight_M_Alpine.Skiing <- list()
# sd_years_weight_M_Alpine.Skiing <- list()
# mean_years_height_M_Alpine.Skiing <- list()
# sd_years_height_M_Alpine.Skiing <- list()
# # women data
# for (i in 1:length(data_Alpine.Skiing_F_year)) {
# 	mean_years_weight_F_Alpine.Skiing[[i]] <- mean(data_Alpine.Skiing_F_year[[i]]$Weight)
# 	sd_years_weight_F_Alpine.Skiing[[i]] <- sd(data_Alpine.Skiing_F_year[[i]]$Weight)
# 	mean_years_height_F_Alpine.Skiing[[i]] <- mean(data_Alpine.Skiing_F_year[[i]]$Height)
# 	sd_years_height_F_Alpine.Skiing[[i]] <- sd(data_Alpine.Skiing_F_year[[i]]$Height)
# }
# # men data
# for (i in 1:length(data_Alpine.Skiing_M_year)) {
# 	mean_years_weight_M_Alpine.Skiing[[i]] <- mean(data_Alpine.Skiing_M_year[[i]]$Weight)
# 	sd_years_weight_M_Alpine.Skiing[[i]] <- sd(data_Alpine.Skiing_M_year[[i]]$Weight)
# 	mean_years_height_M_Alpine.Skiing[[i]] <- mean(data_Alpine.Skiing_M_year[[i]]$Height)
# 	sd_years_height_M_Alpine.Skiing[[i]] <- sd(data_Alpine.Skiing_M_year[[i]]$Height)
# }
# # make data frames
# Alpine.Skiing_F_data <- do.call(rbind, Map(data.frame, Year = years_F_Alpine.Skiing, Mean_Weight = mean_years_weight_F_Alpine.Skiing, StdDev_Weight = sd_years_weight_F_Alpine.Skiing, Mean_Height = mean_years_height_F_Alpine.Skiing, StdDev_Height = sd_years_height_F_Alpine.Skiing))
# Alpine.Skiing_M_data <- do.call(rbind, Map(data.frame, Year = years_M_Alpine.Skiing, Mean_Weight = mean_years_weight_F_Alpine.Skiing, StdDev_Weight = sd_years_weight_F_Alpine.Skiing, Mean_Height = mean_years_height_F_Alpine.Skiing, StdDev_Height = sd_years_height_F_Alpine.Skiing))
```

```{r clean_data}
clean_data <- function(df, sport, sex) {
  sex_list <- rep(sex, nrow(df))
  sport_list <- rep(sport, nrow(df))
  df <- cbind(df, "Sport"=sport_list, "Sex"=sex_list)
  rownames(df) <- c()
  df
}
```


```{r}
# Athletics_F_data <- clean_data(Athletics_F_data, "Athletics", "F")
# Athletics_M_data <- clean_data(Athletics_M_data, "Athletics", "M")
# Swimming_F_data <- clean_data(Swimming_F_data, "Swimming", "F")
# Swimming_M_data <- clean_data(Swimming_M_data, "Swimming", "M")
# Gymnastics_F_data <- clean_data(Gymnastics_F_data, "Gymnastics", "F")
# Gymnastics_M_data <- clean_data(Gymnastics_M_data, "Gymnastics", "M")
# Cross.Country.Skiing_F_data <- clean_data(Cross.Country.Skiing_F_data, "Cross Country Skiing", "F")
# Cross.Country.Skiing_M_data <- clean_data(Cross.Country.Skiing_M_data, "Cross Country Skiing", "M")
# Alpine.Skiing_F_data <- clean_data(Alpine.Skiing_F_data, "Alpine Skiing", "F")
# Alpine.Skiing_M_data <- clean_data(Alpine.Skiing_M_data, "Alpine Skiing", "M")
# 
# head(Athletics_F_data)
# head(Athletics_M_data)
# head(Swimming_F_data)
# head(Swimming_M_data)
# 
# top10_mean_sports <- rbind(Athletics_F_data, Athletics_M_data, Swimming_F_data, Swimming_M_data, Gymnastics_F_data, Gymnastics_M_data, Cross.Country.Skiing_F_data, Cross.Country.Skiing_M_data, Alpine.Skiing_F_data, Alpine.Skiing_M_data)

#write.csv(top10_mean_sports,"/Users/iillari/Documents/gwu/spring2020/intro_dats/model_proj/template/top_5_sports.csv", row.names = F)
```


```{r}
# loadPkg(leaps)
# #This is essentially best fit 
# reg.weight <- regsubsets(Weight ~. , data = less_data, nvmax = ncol(less_data), nbest = 2, method = "exhaustive")  # leaps::regsubsets() - Model selection by exhaustive (default) search, forward or backward stepwise, or sequential replacement
# #The plot will show the Adjust R^2 when using the variables across the bottom
# plot(reg.weight, scale = "r2", main = "R^2")
# plot(reg.weight, scale = "adjr2", main = "Adjusted R^2")
# plot(reg.weight, scale = "cp", main = "Mallows' Cp")
# plot(reg.weight, scale = "bic", main = "Schwartz's information criterion, BIC")
```

```{r}
top10_sports <- data.frame(read.csv("top_10_sports.csv"))
head(top10_sports)
str(top10_sports)
ggplot(top10_sports, aes(x=Year, y=Mean_Weight, color=Sport)) + 
  geom_line() + 
  geom_point() +
  theme_minimal() +
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) + 
  labs(title="Plot of Olympic Weights [kg] Over The Years", x="year", y="weight [kg]") 
```

```{r}
index_F <- which(top10_sports$Sex == "F")
index_M <- which(top10_sports$Sex == "M")
#data_gender <- split(top10_sports, top10_sports$Sex)
top10_sports_F <- top10_sports[index_F,]
# index_Year <- which(top10_sports_F$Year > 1930)
# top10_sports_F <- top10_sports[index_Year,]
top10_sports_M <- top10_sports[index_M,]

plot_top10_F <- ggplot(top10_sports_F, aes(x=Year, y=Mean_Weight, color=Sport)) + 
  geom_line() + 
  geom_point() +
  theme_minimal() +
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) + 
  labs(title="Plot of \nFemale Olympic Weights [kg] \nOver The Years", x="year", y="weight [kg]")
plot_top10_F

plot_top10_M <- ggplot(top10_sports_M, aes(x=Year, y=Mean_Weight, color=Sport)) + 
  geom_line() + 
  geom_point() +
  theme_minimal() +
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) + 
  labs(title="Plot of \nMale Olympic Weights [kg] \nOver The Years", x="year", y="weight [kg]")
plot_top10_M

loadPkg(gridExtra)
grid.arrange(plot_top10_F, plot_top10_M, ncol = 2, nrow = 1)
```


```{r}
head(top10_sports_F)
loadPkg(tidyverse)

medalist <- olympic_data %>% 
  filter(!is.na(Medal)) 

medalist %>% 
  group_by(Medal) %>% 
  summarise(
    mean = mean(Age, na.rm = TRUE)
  ) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover")) 

summer_medalist_sex <-
  medalist %>%
  filter(Season == "Summer") %>% 
  group_by(Year, Sex) %>% 
    summarise(
    mean.Age = mean(Age, na.rm = TRUE)
  )

head(summer_medalist_sex)

summer_ages_gender <- ggplot(data = summer_medalist_sex, 
       mapping = aes(x = Year, y = mean.Age, group = Sex, color = Sex)) +
  geom_point(size=2) +
  geom_line() +
  labs(title="Average Age of Olympic Medalists \nin the Summer Games Separated by Gender",
       x="year", y="mean age [years]") +
  theme_minimal() +
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1))
summer_ages_gender
```

```{r}
filter_data <- function(df, sport) {
  df %>%
    filter(Sport == sport) %>% 
    group_by(Year, Sex) %>%
    summarise(
      mean.Age = mean(Age, na.rm = TRUE),
      sd.Age = sd(Age, na.rm = TRUE),
      mean.Weight = mean(Weight, na.rm = TRUE),
      sd.Weight = sd(Weight, na.rm = TRUE),
      mean.Height = mean(Height, na.rm = TRUE),
      sd.Height = sd(Height, na.rm = TRUE)
  )
}
```

```{r}
make_plots <- function(df, sport, varNameX, varNameY, xVal, yVal, unitX, unitY) {
  var_name_x <- eval(substitute(xVal),eval(df))
  var_name_y <- eval(substitute(yVal),eval(df))
  ggplot(df, mapping = aes(x = var_name_x, y = var_name_y, group = Sex, 
                                        color = Sex)) + geom_point(size=2) + 
    geom_line() +
    labs(title=paste(varNameY, " of Olympic Athletes in ", sport, "\nSeparated by Gender Over ", varNameX),
         x=paste(varNameX, unitX), y=paste(varNameY, unitY)) +
    theme_minimal() +
    theme(panel.border = element_rect(colour = "black", fill=NA, size=1)) 
}
```

## Swimming

```{r}
sports <- olympic_data %>% 
  filter(!is.na(Sport)) 

#head(sports)
swimming <- filter_data(sports, "Swimming")
#head(swimming)

make_plots(swimming, "Swimming", "Years", "Average Age", Year, mean.Age, "[years]","[years]")

make_plots(swimming, "Swimming", "Years", "Average Weight", Year, mean.Weight, "[years]","[kg]")

make_plots(swimming, "Swimming", "Years", "Average Height", Year, mean.Height, "[years]","[cm]")

#grid.arrange(swimming_age, swimming_weight, swimming_height, nrow = 1, ncol = 3)
```

### Models

```{r, results='markup'}
make_model_plots <- function(df, sport, varNameX, varNameY, xVal, yVal, unitX, unitY) {
  var_name_x <- eval(substitute(xVal), eval(df))
  var_name_y <- eval(substitute(yVal), eval(df))
  
  mod_lin <- nlsLM(var_name_y ~ lin_func(var_name_x, a1, b1), 
                   start=list(a1=1, b1=10))
  mod_quad <- nlsLM(var_name_y ~ quad_func(var_name_x, a1, b1, c1), 
                    start=list(a1=1, b1=1, c1=1))
  mod_cubic <- nlsLM(var_name_y ~ cubic_func(var_name_x, a1, b1, c1, d1), 
                     start=list(a1=1, b1=1, c1=1, d1=1))
  mod_exp <- nlsLM(var_name_y ~ exp_func(var_name_x, a1, b1, c1), 
                   start=list(a1=1, b1=0, c1=1))
  
  data_plot <- ggplot(df, aes(x = Year, y=mean.Age, color=Sex)) + 
    geom_point() +
    theme_minimal() +
    labs(title=paste(varNameY, " of Olympic Athletes in ", sport, "\nSeparated by Gender Over ", varNameX), 
         x=paste(varNameX, unitX), y=paste(varNameY, unitY)) + 
    theme(panel.border = element_rect(colour = "black", fill=NA, size=1),
        axis.text.x = element_text(angle = 90, vjust = 0.5)) 
  
  plot_models <- data_plot + geom_line(aes(x=var_name_x, y=predict(mod_lin)), color="red") + 
    geom_line(aes(x=var_name_x, y=predict(mod_quad)), color="orange") +
    geom_line(aes(x=var_name_x, y=predict(mod_cubic)), color="green") +
    geom_line(aes(x=var_name_x, y=predict(mod_exp)), color="blue")
  
  anova_res <- anova(mod_lin, mod_quad, mod_exp, mod_cubic)
  
  return(list(plot_models, anova_res))
}
```


#### Female Athletes

```{r, results='markup'}
index_F <- which(swimming$Sex == "F")
swimming_F <- swimming[index_F, ]

res_Swimming_Age <- make_model_plots(swimming_F, "Swimming", "Years", "Average Age", Year, mean.Age, "[years]","[years]")
res_Swimming_Age[[1]]
res_Swimming_Age[[2]] %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

res_Swimming_Weight <- make_model_plots(swimming_F, "Swimming", "Years", "Average Weight", Year, mean.Weight, "[years]","[kg]")
res_Swimming_Weight[[1]]
res_Swimming_Weight[[2]] %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

res_Swimming_Height <- make_model_plots(swimming_F, "Swimming", "Years", "Average Height", Year, mean.Height, "[years]","[cm]")
res_Swimming_Height[[1]]
res_Swimming_Height[[2]] %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

#### Male Athletes

```{r}
index_M <- which(swimming$Sex == "M")
swimming_M <- swimming[index_M, ]

res_Swimming_Age <- make_model_plots(swimming_M, "Swimming", "Years", "Average Age", Year, mean.Age, "[years]","[years]")
res_Swimming_Age[[1]]
res_Swimming_Age[[2]] %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

res_Swimming_Weight <- make_model_plots(swimming_M, "Swimming", "Years", "Average Weight", Year, mean.Weight, "[years]","[kg]")
res_Swimming_Weight[[1]]
res_Swimming_Weight[[2]] %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

res_Swimming_Height <- make_model_plots(swimming_M, "Swimming", "Years", "Average Height", Year, mean.Height, "[years]","[cm]")
res_Swimming_Height[[1]]
res_Swimming_Height[[2]] %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```
```





## Athletics

```{r}
athletics <- filter_data(sports, "Athletics")

make_plots(athletics, "Athletics", "Years", "Average Age", Year, mean.Age, "[years]","[years]")

make_plots(athletics, "Athletics", "Years", "Average Weight", Year, mean.Weight, "[years]","[kg]")

make_plots(athletics, "Athletics", "Years", "Average Height", Year, mean.Height, "[years]","[cm]")
```

## Gymnastics 

```{r}
gymnastics <- filter_data(sports, "Gymnastics")

make_plots(gymnastics, "Gymnastics", "Years", "Average Age", Year, mean.Age, "[years]","[years]")

make_plots(gymnastics, "Gymnastics", "Years", "Average Weight", Year, mean.Weight, "[years]","[kg]")

make_plots(gymnastics, "Gymnastics", "Years", "Average Height", Year, mean.Height, "[years]","[cm]")
```

## Rowing

```{r}
rowing <- filter_data(sports, "Rowing")

make_plots(rowing, "Rowing", "Years", "Average Age", Year, mean.Age, "[years]","[years]")

make_plots(rowing, "Rowing", "Years", "Average Weight", Year, mean.Weight, "[years]","[kg]")

make_plots(rowing, "Rowing", "Years", "Average Height", Year, mean.Height, "[years]","[cm]")
```

## Baskeball

```{r}
basketball <- filter_data(sports, "Basketball")

make_plots(basketball, "Basketball", "Years", "Average Age", Year, mean.Age, "[years]","[years]")

make_plots(basketball, "Basketball", "Years", "Average Weight", Year, mean.Weight, "[years]","[kg]")

make_plots(basketball, "Basketball", "Years", "Average Height", Year, mean.Height, "[years]","[cm]")
```

## Softball

```{r}
softball <- filter_data(sports, "Softball")

make_plots(softball, "Softball", "Years", "Average Age", Year, mean.Age, "[years]","[years]")

make_plots(softball, "Softball", "Years", "Average Weight", Year, mean.Weight, "[years]","[kg]")

make_plots(softball, "Softball", "Years", "Average Height", Year, mean.Height, "[years]","[cm]")
```

## Fencing

```{r}
fencing <- filter_data(sports, "Fencing")

make_plots(fencing, "Fencing", "Years", "Average Age", Year, mean.Age, "[years]","[years]")

make_plots(fencing, "Fencing", "Years", "Average Weight", Year, mean.Weight, "[years]","[kg]")

make_plots(fencing, "Fencing", "Years", "Average Height", Year, mean.Height, "[years]","[cm]")
```

```{r unload, include=F}
unloadPkg(tidyverse)
unloadPkg(gridExtra)
unloadPkg(reshape2)
unloadPkg(leaps)
unloadPkg(minpack.lm)
unloadPkg(stats)
unloadPkg(kableExtra)
unloadPkg(plyr)
unloadPkg(ggplot2)
```



# References

