---
title: "Olympics_EDA"
author: "LNT"
date: "04/25/2020"
output:
 rmdformats::readthedown:
    highlight: kate
    code_folding: hide
---


```{r basic, include=F}
# use this function to conveniently load libraries and work smoothly with knitting
# can add quietly=T option to the require() function
loadPkg = function(pkg, character.only = FALSE) { 
  if (!character.only) { pkg <- as.character(substitute(pkg)) }
  pkg <- ifelse(!character.only, as.character(substitute(pkg)) , pkg)  
  if (!require(pkg,character.only=T, quietly =T)) {  install.packages(substitute(pkg),dep=T); if(!require(pkg,character.only=T)) stop("Package not found") } 
}
loadPkg(knitr)

# unload/detact package when done using it
unloadPkg = function(pkg, character.only = FALSE) { 
  if(!character.only) { pkg <- as.character(substitute(pkg)) } 
  search_item <- paste("package", pkg,sep = ":") 
  while(search_item %in% search()) { detach(search_item, unload = TRUE, character.only = TRUE) } 
}
```

```{r setup, echo=FALSE, cache=FALSE}


## Global options
options(max.print="75")
knitr::opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
knitr::opts_knit$set(width=75)
options(scientific=T, digits = 3) 
```

```{r xkablesummary}
loadPkg("xtable")
loadPkg("kableExtra")
loadPkg("stringi")

xkabledply = function(smmry, title='Caption', pos='left') { # Thanks Ryan Longmuir for the codes
  smmry %>%
    xtable() %>% 
    kable(caption = title, digits = 4) %>%
    kable_styling(position = "center") %>%
    kable_styling(bootstrap_options = "striped", full_width = F,
    position = pos)
}

xkablesummary = function(df) { 
  #' Combining base::summary, xtable, and kableExtra, to easily display numeric variable summary of dataframes. 
  #` If the categorical variables has less than 6 levels, the function will still run without error.
  #' ELo 202003 GWU DATS
  #' version 1
  #' @param df The dataframe.
  #' @return The summary table for display, or for knitr to process into other formats 
  #' @examples
  #' xkablesummary( faraway::ozone )
  #' xkablesummary( ISLR::Hitters )
  
  s = summary(df) %>%
    apply( 2, function(x) stringr::str_remove_all(x,c("Min.\\s*:\\s*","1st Qu.\\s*:\\s*","Median\\s*:\\s*","Mean\\s*:\\s*","3rd Qu.\\s*:\\s*","Max.\\s*:\\s*")) ) %>% # replace all leading words
    apply( 2, function(x) stringr::str_trim(x, "right")) # trim trailing spaces left
  
  colnames(s) <- stringr::str_trim(colnames(s))
  
  if ( dim(s)[1] ==6 ) { rownames(s) <- c('Min','Q1','Median','Mean','Q3','Max') 
  } else if ( dim(s)[1] ==7 ) { rownames(s) <- c('Min','Q1','Median','Mean','Q3','Max','NA') }
  
  s %>%
    xkabledply("Table: Statistics summary.", "center")

}

xkablevif = function(model) { 
  #' Combining faraway::vif, xtable, and kableExtra, to easily display numeric summary of VIFs for a model. 
  #' ELo 202003 GWU DATS
  #' version 1
  #' @param df The dataframe.
  #' @return The summary table for display, or for knitr to process into other formats 
  #' @examples
  #' xkablevif( model )
  
  vifs = table( names(model$coefficients)[2:length(model$coefficients)] ) # remove intercept to set column names
  vifs[] = faraway::vif(model) # set the values

  vifs %>%
    xtable() %>% 
    kable(caption = "VIFs of the model", digits = 4, col.names = 'VIF') %>% # otherwise it will only has the generic name as 'V1' for the first vector in the table
    kable_styling(position = "center") %>%
    kable_styling(bootstrap_options = "striped", full_width = F,
    position = "left")
}
```

```{r loadpackages}
loadPkg("digest")
loadPkg("corrplot")
loadPkg("ggcorrplot")
loadPkg("ggplot2")
loadPkg("rmdformats")
loadPkg("knitr")
loadPkg("lattice") 
loadPkg("jtools")
loadPkg("faraway")
loadPkg("leaps")
loadPkg("gridExtra")
```

```{r read_data}
# Load data
# 
olympics <- read.csv("olympic_data.csv") 
olympics <- na.omit(olympics)
```


```{r structure}
str(olympics)


```

```{r keep, results = "markup"}

olympics$Sex.Int <- c(as.numeric(as.factor(olympics$Sex)))
olympics$NOC.Int <- c(as.numeric(as.factor(olympics$NOC)))
olympics$Sport.Int <- c(as.numeric(as.factor(olympics$Sport)))

```


#Olympics Correlation plot

```{r data_subset}
loadPkg("dplyr")

olympics_subset <- olympics %>% filter(!is.na(Age)) %>% select(Year, NOC.Int, Sex.Int, Age, Height, Weight, BMI, BMI.Category, Population, GDP, GDPpC, Medal.No.Yes)


unloadPkg("dplyr")
```


```{r pairs_logit1}
#loadPkg(psych) # pair plots with histogram on diagonal and other options
#pairs.panels(olympics_subset, 
#             method = "pearson", # correlation method
#             hist.col = "#00AFBB",
#             density = FALSE,  # show density plots
#             ellipses = FALSE # show correlation ellipses
#             )
#unloadPkg(psych)
```

```{r}


olympics_cor <- cor(olympics_subset,use="complete.obs")

corrplot.mixed(olympics_cor)


```

```{r Tmedals}
#loadPkg("dplyr")

#olympics_medal_count <- olympics_subset %>% group_by(Year) %>% tally()

#unloadPkg("dplyr")

```

# Olympics 1915-1925.  First World War (1914-18) and 1918 (H1N1) Pandemic (aka Spanish Flu)
[https://www.ncbi.nlm.nih.gov/books/NBK22148/] (https://www.ncbi.nlm.nih.gov/books/NBK22148/)
The following countries in Europe had 2.64 million excess deaths occurred during the period when the H1N1 Pandemic (Spanish Flu) was circulating from January 1918 - June 1919: Italy, Bulgaria, Portugal, Spain, Netherlands, Sweden, Germany, Switzerland, France, Norway, Denmark, UK (Scotland, England, Wales). In the US, 675,000 people died from H1N1 which nearly was 0.8 percent of the 1910 population. 


```{r read_pandemic_data}

p_olympics <- read.csv("pandemic_olympics.csv")


```
(JOHNSON, NIALL P. A. S., and JUERGEN MUELLER. “Updating the Accounts: Global Mortality of the 1918-1920 ‘Spanish’ Influenza Pandemic.” Bulletin of the History of Medicine, vol. 76, no. 1, 2002, pp. 105–115. JSTOR, www.jstor.org/stable/44446153. Accessed 19 Apr. 2020.)

```{r H1N1_pandemic}
loadPkg("dplyr")
NOC_SF <- c("ITA", "NED", "SWE", "FRA", "NOR", "DEN", "GBR", "USA")
Medals <- c("Gold", "Silver", "Bronze")

pandemic_NOC_Yr_Mdl <- p_olympics %>% filter(Year >= 1908 & Year <= 1928,
                                         NOC %in% NOC_SF,
                                         Medal %in% Medals) %>% group_by(NOC, Year) %>% tally()

pandemic_NOC_Yr_Mdl$Year <- as.factor(pandemic_NOC_Yr_Mdl$Year)
unloadPkg("dplyr")

```


```{r plots_num_medals , include=TRUE}
NOC_colors <- c("#A6CEE3", "#1F78B4", "#B2DF8A", "#33A02C", "#FB9A99", "#E31A1C", "#FDBF6F","#FF7F00")


pandemic_df <- data.frame(pandemic_NOC_Yr_Mdl$NOC, pandemic_NOC_Yr_Mdl$Year, pandemic_NOC_Yr_Mdl$n)
pandemic_df <- unique(pandemic_df)
colnames(pandemic_df) <- c("NOC", "Year", "Total.Medals")

pandemic_df$NOC <- as.factor(pandemic_df$NOC)



# Basic line plot with points
ggplot(data=pandemic_df, aes(x=Year, y=Total.Medals, group=NOC, color=NOC)) +
  geom_line(size=1)+
#  scale_colour_manual(values = NOC_colors) +
   geom_vline(xintercept = 1918) +
  geom_point()+
 theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, size = 9, hjust = 1),
        axis.text.y = element_text(vjust = 1, size = 9),
        panel.border = element_rect(color = "black", fill=NA, size=1)) +
  labs(title="H1N1 Pademic: Num of medals (n) vs. Year (1908-1928)", x="Year", y="Number of medals (n)") 


```

```{r plots_num_athletes, include=TRUE}
loadPkg("dplyr")
NOC_SF <- c("ITA", "NED", "SWE", "FRA", "NOR", "DEN", "GBR", "USA")

p_num_athletes <- p_olympics %>% filter(Year >= 1908 & Year <= 1928,
                                         NOC %in% NOC_SF) %>% group_by(NOC, Year) %>% tally(ID)


p_num_athletes_df <- data.frame(p_num_athletes$NOC, p_num_athletes$Year, p_num_athletes$n)
p_num_athletes_df <- unique(p_num_athletes_df)
colnames(p_num_athletes_df) <- c("NOC", "Year", "Total.Athletes")
p_num_athletes_df$NOC <- as.factor(p_num_athletes_df$NOC)
p_num_athletes_df$Year <- as.factor(p_num_athletes_df$Year)
p_num_athletes_df$V_line <- c("1918") 


# Basic line plot with points
ggplot(data=p_num_athletes_df, aes(x=Year, y=Total.Athletes, group=NOC, color=NOC)) +
  geom_line(size=1)+
#  scale_colour_manual(values = NOC_colors) +
   geom_vline(xintercept = 1918) +
  geom_point()+
 theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, size = 9, hjust = 1),
        axis.text.y = element_text(vjust = 1, size = 9),
        panel.border = element_rect(color = "black", fill=NA, size=1)) +
  labs(title="H1N1 Pademic: Num of athletes (n) vs. Year (1908-1928)", x="Year", y="Number of Athletes (n)")
  
unloadPkg("dplyr")
```


```{r ageSF}
loadPkg(dplyr)
NOC_colors <- c("#A6CEE3", "#1F78B4", "#B2DF8A", "#33A02C", "#FB9A99", "#E31A1C", "#FDBF6F","#FF7F00")

pandemic_avg_age <- p_olympics %>% filter(Year >= 1908 & Year <= 1928,
                                         NOC %in% NOC_SF,
                                         !is.na(Age)) %>% group_by(NOC, Year) %>% summarise(avg=mean(Age))

pandemic_avg_age$Year <- as.factor(pandemic_avg_age$Year)


# Basic line plot with points
ggplot(data=pandemic_avg_age, aes(x=Year, y=avg, group=NOC, color=NOC)) +
  geom_line(size=1)+
   geom_vline(xintercept = 1918) +
  geom_point()+
 theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, size = 9, hjust = 1),
        axis.text.y = element_text(vjust = 1, size = 9),
        panel.border = element_rect(color = "black", fill=NA, size=1)) +
  labs(title="H1N1 Pademic: Average Age vs. Year (1908-1928)", x="Year", y="Number of Athletes (n)")
```


```{r w_h}
pandemic_avg_hw<- p_olympics %>% filter(Year >= 1908 & Year <= 1928,
                                         NOC %in% NOC_SF,
                                        !is.na(Height),
                                         !is.na(Weight)) %>% group_by(NOC, Year) %>% summarise(avg_h = mean(Height), avg_w = mean(Weight))
pandemic_avg_hw$Year <- as.factor(pandemic_avg_hw$Year)


# Basic line plot with points
h1 <- ggplot(data=pandemic_avg_hw, aes(x=Year, y=avg_h, group=NOC, color=NOC)) +
  geom_line(size=1)+
   geom_vline(xintercept = 1918) +
  geom_point()+
 theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, size = 9, hjust = 1),
        axis.text.y = element_text(vjust = 1, size = 9),
        panel.border = element_rect(color = "black", fill=NA, size=1)) +
  labs(title="H1N1 Pademic: Average Height vs. Year (1908-1928)", x="Year", y="Avg Height (cm)")
h1

w1 <- ggplot(data=pandemic_avg_hw, aes(x=Year, y=avg_w, group=NOC, color=NOC)) +
  geom_line(size=1)+
   geom_vline(xintercept = 1918) +
  geom_point()+
 theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, size = 9, hjust = 1),
        axis.text.y = element_text(vjust = 1, size = 9),
        panel.border = element_rect(color = "black", fill=NA, size=1)) +
  labs(title="H1N1 Pademic: Average Weight vs. Year (1908-1928)", x="Year", y="Avg Height (cm)")
w1

```

# Model Building

```{r model1}
loadPkg("dplyr")
olympics_medal_count <- olympics_subset %>% group_by(Year) %>% tally(Medal.No.Yes)

p_olympics$Medal.No.Yes <- ifelse(p_olympics$Medal == "Gold" | p_olympics$Medal == "Silver" | p_olympics$Medal == "Bronze",1,0)
p_olympics$Medal.No.Yes[is.na(p_olympics$Medal.No.Yes)] <- 0

#p_olympics$Medal.No.Yes <- as.factor(p_olympics$Medal.No.Yes)


pandemic_total_medals_gold<- p_olympics %>% filter(Medal == "Gold", Season == "Summer") %>% group_by(NOC, Year) %>% tally(Medal.No.Yes)
#pandemic_total_medals_gold
pandemic_total_medals_silver<- p_olympics %>% filter(Medal == "Silver", Season == "Summer") %>% group_by(NOC, Year) %>% tally(Medal.No.Yes)
#pandemic_total_medals_silver
pandemic_total_medals_bronze<- p_olympics %>% filter(Medal == "Bronze", Season == "Summer") %>% group_by(NOC, Year) %>% tally(Medal.No.Yes)
#pandemic_total_medals_bronze
pandemic_total_medals <- p_olympics %>% filter(Medal %in% Medals, Season == "Summer") %>% group_by(NOC, Year) %>% tally(Medal.No.Yes)
#pandemic_total_medals

df<- data.frame(p_olympics$Medal, p_olympics$Medal.No.Yes)
df

unloadPkg("dplyr")
```


```{r }
loadPkg("dplyr")

olympics_total_medals <- olympics %>% filter(Medal %in% Medals, Season == "Summer") %>%
group_by(NOC, Year) %>% tally(Medal.No.Yes)
#olympics_total_medals



ITA_Tmedals <- pandemic_total_medals %>% filter(NOC == "ITA")  
#ITA_Tmedals
ITA_Tmedals_all <- olympics_total_medals  %>% filter(NOC == "ITA")  
#ITA_Tmedals_all


#histogram of Number of Medals
loadPkg("ggplot2")
ggplot(olympics_total_medals, aes(x=Year))  + geom_histogram(color="gray7", aes(y=..density..)) + ylab("Relative Frequency") 
#unloadPkg("ggplot2")
unloadPkg("dplyr")
```

```{r forecast}

olympic_pts <- ts(ITA_Tmedals$n, start=1908, end = 1928, deltat=4)
olympic_all <- ts(ITA_Tmedals_all$n, start=1960, end = 2016, deltat=4)

ita_olympic <- ts.union(olympic_pts, olympic_all)
plot(ita_olympic)



loadPkg("tidyverse")
ita_df <- data.frame(ita_olympic)
ita_df <- ita_df %>% mutate(Num.Medals= if_else(is.na(ita_df$olympic_pts), ita_df$olympic_all,ita_df$olympic_pts ))

ita_df <- subset(ita_df, select = c(Num.Medals))

ita_df

ol_ts <- ts(ita_df,start=1908, end=2016, deltat=4 )
ol_ts
plot(ol_ts)

olympic_train <- window(ol_ts, start=1908, end = 1928, deltat=4)

olympic_test <- window(ol_ts, start=2012, end = 2016, deltat=4)

```
```{r forecast1}

olympic_pts <- ts(ITA_Tmedals$n, start=1908, end = 1928, deltat=4)
olympic_all <- ts(ITA_Tmedals_all$n, start=1960, end = 2016, deltat=4)

olympic_train <- window(olympic_pts, start=1908, end = 1928, deltat=4)
olympic_train1 <- window(ol_ts, start=1960, end = 2012, deltat=4)
olympic_test <- window(ol_ts, start=2012, end = 2016, deltat=4)

```

```{r fit}
fit1.hw <- HoltWinters(x=olympic_train, beta = FALSE,gamma = FALSE) 
plot(fit1.hw)
summary(fit1.hw)

fit2.hw <- HoltWinters(x=olympic_train1, beta = FALSE, gamma = FALSE) 
plot(fit2.hw)
summary(fit2.hw)

loadPkg("forecast")
fit1.ets <-ets(olympic_train1, model="ANN")
plot(fit1.ets)
fit1_forecast <- forecast(fit1.ets, h = 3) 
fit1_forecast

fit1.hw.forecast <- forecast(fit1.hw, 8)
plot(fit1.hw.forecast)

fit2.hw.forecast <- forecast(fit2.hw, 8)
plot(fit2.hw.forecast)

```
```{r }
h <- 10
fit.lin <- tslm(olympic_all ~ trend)
fcasts.lin <- forecast(fit.lin, h = h)
fit.exp <- tslm(olympic_all ~ trend, lambda = 0)
fcasts.exp <- forecast(fit.exp, h = h)

t <- time(olympic_all)
t.break1 <- 1960
t.break2 <- 2016
tb1 <- ts(pmax(0, t - t.break1), start = 1960)
tb2 <- ts(pmax(0, t - t.break2), start = 1960)

fit.pw <- tslm(olympic_all ~ t + tb1 + tb2)
t.new <- t[length(t)] + seq(h)
tb1.new <- tb1[length(tb1)] + seq(h)
tb2.new <- tb2[length(tb2)] + seq(h)

newdata <- cbind(t=t.new, tb1=tb1.new, tb2=tb2.new) %>%
  as.data.frame()
fcasts.pw <- forecast(fit.pw, newdata = newdata)

fit.spline <- tslm(olympic_all ~ t + I(t^2) + I(t^3) +
  I(tb1^3) + I(tb2^3))
fcasts.spl <- forecast(fit.spline, newdata = newdata)

autoplot(olympic_all) +
  autolayer(fitted(fit.lin), series = "Linear") +
  autolayer(fitted(fit.exp), series = "Exponential") +
  autolayer(fitted(fit.pw), series = "Piecewise") +
  autolayer(fitted(fit.spline), series = "Cubic Spline") +
  autolayer(fcasts.pw, series="Piecewise") +
  autolayer(fcasts.lin, series="Linear", PI=FALSE) +
  autolayer(fcasts.exp, series="Exponential", PI=FALSE) +
  autolayer(fcasts.spl, series="Cubic Spline", PI=FALSE) +
  xlab("Year") + ylab("Number of Medals") +
  ggtitle("Italy Olympics") +
  guides(colour = guide_legend(title = " "))

```

```{r func1}
# linear model function, it takes 3 inputs
lin_func <- function(x, a, b) {
  a*x+b
}
# quadratic model function, it takes 4 inputs
quad_func <- function(x, a, b, c) {
  a*x*x + b*x + c
}
# cubic model function, it takes 5 inputs
cubic_func <- function(x, a, b, c, d) {
  a*x*x*x + b*x*x + c*x + d 
}
# exponential model function, it takes 4 inputs
exp_func <- function(x, a, b, c) {
  a*exp(b*x) + c
}
```


```{r}
loadPkg("tidyverse")
loadPkg("minpack.lm")
xdata <- olympics_total_medals$Year
ydata <- olympics_total_medals$n
mod_lin <- nlsLM(ydata~lin_func(xdata, a1, b1), start=list(a1=1, b1=20))
mod_quad <- nlsLM(ydata~quad_func(xdata, a1, b1, c1), start=list(a1=1, b1=1, c1=1))
mod_cubic <- nlsLM(ydata~cubic_func(xdata, a1, b1, c1, d1), start=list(a1=1, b1=1, c1=1, d1=1))
mod_exp <- nlsLM(ydata~exp_func(xdata, a1, b1, c1), start=list(a1=1, b1=0, c1=1))
data_plot <- ggplot(olympics_total_medals, aes(x = Year, y=n)) + 
  geom_point() + 
  ylim(0, 80) + xlim(1960, 2020) +
  theme_minimal() +
  labs(title="Plot of the number of Medals at the Olympic Game over the years \n starting from 1948", x="Year", y="Number of Medals") + 
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1),
        axis.text.x = element_text(angle = 90, vjust = 0.5)) 
data_plot + geom_line(aes(x=xdata, y=predict(mod_lin)), color="red") + 
  geom_line(aes(x=xdata, y=predict(mod_quad)), color="orange") +
  geom_line(aes(x=xdata, y=predict(mod_cubic)), color="green") +
  geom_line(aes(x=xdata, y=predict(mod_exp)), color="blue")
model_list_1exp <- c("Linear", "Exponential")
model_list_123 <- c("Linear", "Quadratic", "Cubic")
anova_results1exp <- anova(mod_lin, mod_exp)
anova_results123 <- anova(mod_lin, mod_quad, mod_cubic) 
anova_res1exp <- cbind("Models"=model_list_1exp, anova_results1exp)
anova_res123 <- cbind("Models"=model_list_123, anova_results123)
anova_res1exp %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
anova_res123 %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```


```{r arima}
olympic_arima<- auto.arima(olympic_all, seasonal=FALSE)
summary(olympic_arima)

olympic_arima %>% forecast(h=10) %>% autoplot(include=80)

```

## Linear Model
#Create New_data Adding Total Number of Medals
```{r}
loadPkg("dplyr")
new_data <- olympics %>% filter(Medal.No.Yes == "1") %>% group_by(Year) %>% add_count(Medal, name="TMedals")


unloadPkg("dplyr")

```

```{r, results='asis'}
xkablesummary(new_data)
```

#Create Correlation Plot with New_data
```{r corr}
loadPkg("GGally")
ggcorr(new_data, label= TRUE, label_alpha = TRUE)
unloadPkg("GGally")
```


## Linear Model1 TMedals ~ BMI
```{r linear_model, results="markup"}
model1 = lm(TMedals ~ BMI, data=new_data)
sum_md1 = summary(model1)
sum_md1

```

```{r vif_md1}

vif_md1 = faraway::vif(model1)
```

```{r, results='asis'}
xkabledply(sum_md1, "Linear model summary", "center")

xkablevif(model1)
```

## Linear Model2 TMedals ~ Age
```{r model2}
model2 = lm(TMedals ~ Age, data=new_data)
sum_md2 = summary(model2)
sum_md2
```

```{r vifmodel2}
vif_md2 = faraway::vif(model2)
vif_md2
```



```{r, results='asis'}
xkabledply(sum_md2, "Linear model2 summary", "center")

xkablevif(model2)
```

## Linear Model 3 TMedals ~ BMI + Age
```{r model3}
model3 = lm(TMedals ~ BMI + Age , data=new_data)
sum_md3 = summary(model3)
sum_md3
```

```{r, results='asis'}
xkabledply(sum_md3, "Linear model3 summary", "center")

xkablevif(model3)
```

## Linear Model 4 TMedals ~ BMI + Age + GDP
```{r model4}
model4 = lm(TMedals ~ BMI + Age + GDP , data=new_data)
sum_md4 = summary(model4)
sum_md4

```

```{r, results='asis'}
xkabledply(sum_md4, "Linear model4 summary", "center")

xkablevif(model4)
```

```{r q04_summary, echo=TRUE}
loadPkg(kableExtra)
loadPkg(jtools)

model_names <- c("Linear Model 1: TMedals ~ BMI", "Linear Model 2: TMedals ~ Age", "Linear Model 3: TMedals ~ BMI + Age", "Linear Model 4: TMedals ~ BMI + Age+ GDP")
m <- c(summary(model1))
ar2_values <- c(summary(model1)$adj.r.squared,summary(model2)$adj.r.squared, summary(model3)$adj.r.squared, summary(model4)$adj.r.squared)
df <- data.frame(model_names, ar2_values)

names(df) <- c("", "Adjusted R2")

kable(df) %>% kable_styling()

vdf <- data.frame(vif_md2)
  
vtable<-  kable(vdf) %>% kable_styling




```

```{r anova}
#anova_m <- anova(model1, model2, model3, model4)
#summary(anova_m)
```

```{r corr1}

new_data1 <- olympics  %>% group_by(Year) %>% add_count(Medal, name="TMedals")

keep_columns <- c( "NOC", "Year", "Decade", "Sex", "Age", "Height", "Weight", "BMI", "Population", "GDP", "Medal.No.Yes", "TMedals")


new_data_subset <- new_data1[keep_columns]
new_data_subset <- new_data_subset %>% filter(!is.na(Age))
```

```{r pairs_logit}
loadPkg(psych) # pair plots with histogram on diagonal and other options
pairs.panels(new_data_subset , 
             method = "pearson", # correlation method
             hist.col = "#00AFBB",
             density = FALSE,  # show density plots
             ellipses = FALSE # show correlation ellipses
             )
unloadPkg(psych)

```
```{r ggplot}
loadPkg(ggplot2)
new_data_subset$Medal.No.Yes=as.factor(new_data_subset$Medal.No.Yes)


ggplot(new_data_subset, aes(Age, fill = Medal.No.Yes)) + geom_density(alpha = 0.2)

ggplot(new_data_subset, aes(BMI, fill = Medal.No.Yes)) + geom_density(alpha = 0.2)

```

```{r logit}

olympic_logit <- glm(Medal.No.Yes ~ Age + Height+ Weight, data = new_data_subset, binomial(link = "logit"))
summary(olympic_logit)
exp(coef(olympic_logit))



## CIs using standard errors
confint.default(olympic_logit)
```






```{r HosmerLemeshow, eval=FALSE}
#### Hosmer and Lemeshow test  
########The Hosmer and Lemeshow Goodness of Fit test can be used to evaluate logistic regression fit. 

loadPkg(ResourceSelection) # function hoslem.test( ) for logit model evaluation
olympic_logit1_Hoslem = hoslem.test(new_data_subset$Medal.No.Yes, fitted(olympic_logit)) # Hosmer and Lemeshow test, a chi-squared test
unloadPkg(ResourceSelection) 
olympic_logit1_Hoslem

```

```{r AUC}

loadPkg(pROC) # receiver operating characteristic curve, gives the diagnostic ability of a binary classifier system as its discrimination threshold is varied. The curve is on sensitivity/recall/true-positive-rate vs false_alarm/false-positive-rate/fall-out.
new_data_subset$prob=predict(olympic_logit, type = c("response")) # Add new column of predicted probabilities
h <- roc(Medal.No.Yes~prob, data=new_data_subset)
auc(h) # area-under-curve prefer 0.8 or higher.
plot(h)


```
```{r confusion}
loadPkg(regclass)
olympic_logit_confusion = confusion_matrix(olympic_logit)
unloadPkg(regclass)
```

```{r McFadden}
olympic_nulllogit2 <- glm(Medal.No.Yes ~ 1, data = new_data_subset, family = "binomial")
mcFadden = 1 - logLik(olympic_logit)/logLik(olympic_nulllogit2)
mcFadden


```